---
title: "AGS paper I - Supplementary Information (SI)"
author: "[John Zobolas](https://github.com/bblodfon)"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
description: "AGS paper I - SI"
url: 'https\://username.github.io/reponame/'
github-repo: "username/reponame"
bibliography: ["references.bib", "packages.bib"]
link-citations: true
site: bookdown::bookdown_site
---

```{r include = FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(.packages(), 'emba', 'usefun', 'rtemps', 'PRROC', 'glmnet', 'MAMSE'), 'packages.bib')
```

# Intro {-}

This report is the **supplementary material** for the AGS I Paper and has all the simulation results and investigations related to that paper, as well as guidelines for reproducing the results.

A list of things that change between the simulations and the presented graphs are:

- The number of `Gitsbe` simulations: more simulations, more models generated.
- The type of mutation that `Gitsbe` models have:
unless otherwise specified, the `Gitsbe` models have only [link operator mutations](https://druglogics.github.io/druglogics-doc/gitsbe-config.html#genetic-algorithm).
[Topology mutations] were also tested as well as a combination of [topology and link operator mutations].
- The [training data](https://druglogics.github.io/druglogics-doc/training-data.html) for the `Gitsbe` models: *steady state* (calibrated models) vs *proliferative profile* (proliferative models).
Also *randomly generated models* were produced for the link operator mutations using the [abmlog module](https://github.com/druglogics/abmlog).
- The type of mathematical model (HSA or Bliss) used in `Drabme` to evaluate the synergies either from the [@Flobak2015] for the CASCADE 1.0 analysis or from the [@Flobak2019] dataset for the CASCADE 2.0 analysis.
More info on the calcualtions that Drabme does [see here](https://druglogics.github.io/druglogics-doc/drabme-description.html#drabme-description).
- The type of output used from `Drabme`: ensemble-wise or model-wise [synergy results](https://druglogics.github.io/druglogics-doc/drabme-install.html#drabme-output).

For the ROC curves we used the function `get_roc_stats()` from [@R-usefun] and for the PR curves the `pr.curve()` from [@R-PRROC] (see also [@Grau2015]).

The report template is from the `rtemps` R package [@R-rtemps].

# Input {-}

Loading libraries:
```{r Load libraries, message = FALSE}
library(DT)
library(ggpubr)
library(RColorBrewer)
library(xfun)
library(dplyr)
library(tidyr)
library(tibble)
library(emba)
library(usefun)
library(readr)
library(stringr)
library(latex2exp)
library(corrplot)
library(PRROC)
library(equatiomatic)
library(glmnet)
library(knitr)
library(MAMSE)
```

# CASCADE 1.0 Analysis {-}

:::{.blue-box}
Performance of automatically parameterized models against published data in [@Flobak2015]
:::

## HSA results {-}

:::{.note}
- *HSA* refers to the synergy method used in `Drabme` to assess the synergies from the `gitsbe` models
- We test performance using ROC and PR AUC for both the *ensemble-wise* and *model-wise* synergies from `Drabme`
- **Calibrated** models: fitted to steady state ($50$ simulations)
- **Proliferative** models: fitted to [proliferation profile](https://druglogics.github.io/druglogics-doc/training-data.html#unperturbed-condition---globaloutput-response) ($50$ simulations)
- **Random** models: produced via `abmlog` (see [here](#random-model-results) and used in `Drabme` with `synergy_method: hsa`
- `Gitsbe` models have mutations on **link operator** only
:::

Load results:
```{r Load CASCADE 1.0 HSA results}
# 'ss' => calibrated models, 'random' => random models, 'prolif' => proliferative models
# 'ew' => ensemble-wise, 'mw' => model-wise

## HSA results
ss_hsa_ew_file = paste0("results/link-only/hsa/cascade_1.0_ss_50sim_fixpoints_ensemblewise_synergies.tab")
ss_hsa_mw_file = paste0("results/link-only/hsa/cascade_1.0_ss_50sim_fixpoints_modelwise_synergies.tab")
prolif_hsa_ew_file = paste0("results/link-only/hsa/cascade_1.0_rand_50sim_fixpoints_ensemblewise_synergies.tab")
prolif_hsa_mw_file = paste0("results/link-only/hsa/cascade_1.0_rand_50sim_fixpoints_modelwise_synergies.tab")
random_hsa_ew_file = paste0("results/link-only/hsa/cascade_1.0_random_ensemblewise_synergies.tab")
random_hsa_mw_file = paste0("results/link-only/hsa/cascade_1.0_random_modelwise_synergies.tab")

ss_hsa_ensemblewise_synergies = emba::get_synergy_scores(ss_hsa_ew_file)
ss_hsa_modelwise_synergies = emba::get_synergy_scores(ss_hsa_mw_file, file_type = "modelwise")
prolif_hsa_ensemblewise_synergies = emba::get_synergy_scores(prolif_hsa_ew_file)
prolif_hsa_modelwise_synergies = emba::get_synergy_scores(prolif_hsa_mw_file, file_type = "modelwise")
random_hsa_ensemblewise_synergies = emba::get_synergy_scores(random_hsa_ew_file)
random_hsa_modelwise_synergies = emba::get_synergy_scores(random_hsa_mw_file, file_type = "modelwise")

# calculate probability of synergy in the modelwise results
ss_hsa_modelwise_synergies = ss_hsa_modelwise_synergies %>% 
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
prolif_hsa_modelwise_synergies = prolif_hsa_modelwise_synergies %>%
  mutate(synergy_prob_prolif = synergies/(synergies + `non-synergies`))
random_hsa_modelwise_synergies = random_hsa_modelwise_synergies %>%
  mutate(synergy_prob_random = synergies/(synergies + `non-synergies`))

observed_synergies_file = paste0("results/observed_synergies_cascade_1.0")
observed_synergies = get_observed_synergies(observed_synergies_file)
# 1 (positive/observed synergy) or 0 (negative/not observed) for all tested drug combinations
observed = sapply(random_hsa_modelwise_synergies$perturbation %in% observed_synergies, as.integer)

# 'ew' => ensemble-wise, 'mw' => model-wise
pred_ew_hsa = bind_cols(ss_hsa_ensemblewise_synergies %>% rename(ss_score = score),
  prolif_hsa_ensemblewise_synergies %>% select(score) %>% rename(prolif_score = score),
  random_hsa_ensemblewise_synergies %>% select(score) %>% rename(random_score = score), 
  as_tibble_col(observed, column_name = "observed"))

pred_mw_hsa = bind_cols(
  ss_hsa_modelwise_synergies %>% select(perturbation, synergy_prob_ss),
  prolif_hsa_modelwise_synergies %>% select(synergy_prob_prolif),
  random_hsa_modelwise_synergies %>% select(synergy_prob_random),
  as_tibble_col(observed, column_name = "observed"))
```

### ROC curves {-}

```{r roc-hsa-cascade1, fig.width=5, fig.height=5, dpi=300, cache=TRUE, fig.show='hold', out.width='50%', fig.cap='ROC curves (CASCADE 1.0, HSA synergy method)', fig.align='center'}
res_ss_ew = get_roc_stats(df = pred_ew_hsa, pred_col = "ss_score", label_col = "observed")
res_prolif_ew = get_roc_stats(df = pred_ew_hsa, pred_col = "prolif_score", label_col = "observed")
res_random_ew = get_roc_stats(df = pred_ew_hsa, pred_col = "random_score", label_col = "observed")

res_ss_mw = get_roc_stats(df = pred_mw_hsa, pred_col = "synergy_prob_ss", label_col = "observed", direction = ">")
res_prolif_mw = get_roc_stats(df = pred_mw_hsa, pred_col = "synergy_prob_prolif", label_col = "observed", direction = ">")
res_random_mw = get_roc_stats(df = pred_mw_hsa, pred_col = "synergy_prob_random", label_col = "observed", direction = ">")

# Plot ROCs
my_palette = RColorBrewer::brewer.pal(n = 9, name = "Set1")

plot(x = res_ss_ew$roc_stats$FPR, y = res_ss_ew$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Ensemble-wise synergies (HSA)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = res_prolif_ew$roc_stats$FPR, y = res_prolif_ew$roc_stats$TPR, 
  lwd = 3, col = my_palette[2])
lines(x = res_random_ew$roc_stats$FPR, y = res_random_ew$roc_stats$TPR, 
  lwd = 3, col = my_palette[3])
legend('bottomright', title = 'AUC', col = my_palette[1:3], pch = 19,
  legend = c(paste(round(res_ss_ew$AUC, digits = 3), "Calibrated"), 
    paste(round(res_prolif_ew$AUC, digits = 3), "Proliferative"),
    paste(round(res_random_ew$AUC, digits = 3), "Random")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)

plot(x = res_ss_mw$roc_stats$FPR, y = res_ss_mw$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Model-wise synergies (HSA)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = res_prolif_mw$roc_stats$FPR, y = res_prolif_mw$roc_stats$TPR, 
  lwd = 3, col = my_palette[2])
lines(x = res_random_mw$roc_stats$FPR, y = res_random_mw$roc_stats$TPR, 
  lwd = 3, col = my_palette[3])
legend('bottomright', title = 'AUC', col = my_palette[1:3], pch = 19,
  legend = c(paste(round(res_ss_mw$AUC, digits = 3), "Calibrated"),
    paste(round(res_prolif_mw$AUC, digits = 3), "Proliferative"), 
    paste(round(res_random_mw$AUC, digits = 3), "Random")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)
```

### PR curves {-}

```{r pr-hsa-cascade1, fig.width=5, fig.height=5, dpi=300, cache=TRUE, fig.show='hold', out.width='50%', fig.cap='PR curves (CASCADE 1.0, HSA synergy method)', fig.align='center'}
pr_ss_ew_hsa = pr.curve(scores.class0 = pred_ew_hsa %>% pull(ss_score) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_hsa %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_prolif_ew_hsa = pr.curve(scores.class0 = pred_ew_hsa %>% pull(prolif_score) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_hsa %>% pull(observed), curve = TRUE)
pr_random_ew_hsa = pr.curve(scores.class0 = pred_ew_hsa %>% pull(random_score) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_hsa %>% pull(observed), curve = TRUE)

pr_ss_mw_hsa = pr.curve(scores.class0 = pred_mw_hsa %>% pull(synergy_prob_ss), 
  weights.class0 = pred_mw_hsa %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_prolif_mw_hsa = pr.curve(scores.class0 = pred_mw_hsa %>% pull(synergy_prob_prolif), 
  weights.class0 = pred_mw_hsa %>% pull(observed), curve = TRUE)
pr_random_mw_hsa = pr.curve(scores.class0 = pred_mw_hsa %>% pull(synergy_prob_random), 
  weights.class0 = pred_mw_hsa %>% pull(observed), curve = TRUE)

plot(pr_ss_ew_hsa, main = 'PR curve, Ensemble-wise synergies (HSA)', 
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_prolif_ew_hsa, add = TRUE, color = my_palette[2])
plot(pr_random_ew_hsa, add = TRUE, color = my_palette[3])
legend('topright', title = 'AUC', col = my_palette[1:3], pch = 19,
  legend = c(paste(round(pr_ss_ew_hsa$auc.davis.goadrich, digits = 3), "Calibrated"), 
    paste(round(pr_prolif_ew_hsa$auc.davis.goadrich, digits = 3), "Proliferative"),
    paste(round(pr_random_ew_hsa$auc.davis.goadrich, digits = 3), "Random")))
grid(lwd = 0.5)

plot(pr_ss_mw_hsa, main = 'PR curve, Model-wise synergies (HSA)', 
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_prolif_mw_hsa, add = TRUE, color = my_palette[2])
plot(pr_random_mw_hsa, add = TRUE, color = my_palette[3])
legend('left', title = 'AUC', col = my_palette[1:3], pch = 19,
  legend = c(paste(round(pr_ss_mw_hsa$auc.davis.goadrich, digits = 3), "Calibrated"),
    paste(round(pr_prolif_mw_hsa$auc.davis.goadrich, digits = 3), "Proliferative"),
    paste(round(pr_random_mw_hsa$auc.davis.goadrich, digits = 3), "Random")))
grid(lwd = 0.5)
```

:::{.green-box}
- Calibrated models perform a lot better than either the random or proliferative ones
:::

### AUC sensitivity {-}

:::{#auc-sensitivity .blue-box}
- Investigate **combining the synergy results of calibrated and proliferative (random) models**
- Quantify the amount of information from the *proliferative* (*random*) models that can be used to augment the calibrated results?
- **Ensemble-wise** scenario: $score = calibrated + \beta \times proliferative$ ($random$)
  - $\beta \rightarrow +\infty$: mostly *proliferative* (random) model predictions
  - $\beta \rightarrow -\infty$: mostly *reverse proliferative* (random) model predictions
  - $\beta \simeq -1$: calibrated models are *normalized* against proliferative (random) model predictions
- **Model-wise** scenario: $(1-w) \times prob_{cal} + w \times prob_{rand}, w \in[0,1]$
  - $w=0$: only calibrated model predictions
  - $w=1$: only proliferative (random) model predictions
:::

```{r auc-sen-ew-hsa-cascade1, dpi=300, fig.align='center', cache=TRUE, fig.cap='AUC sensitivity (CASCADE 1.0, HSA synergy method, Ensemble-wise results)'}
# Ensemble-wise
betas = seq(from = -20, to = 20, by = 0.1)

random_roc = sapply(betas, function(beta) {
  pred_ew_hsa = pred_ew_hsa %>% mutate(combined_score = ss_score + beta * random_score)
  res = roc.curve(scores.class0 = pred_ew_hsa %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_hsa %>% pull(observed))
  auc_value = res$auc
})

prolif_roc = sapply(betas, function(beta) {
  pred_ew_hsa = pred_ew_hsa %>% mutate(combined_score = ss_score + beta * prolif_score)
  res = roc.curve(scores.class0 = pred_ew_hsa %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_hsa %>% pull(observed))
  auc_value = res$auc
})

random_pr = sapply(betas, function(beta) {
  pred_ew_hsa = pred_ew_hsa %>% mutate(combined_score = ss_score + beta * random_score)
  res = pr.curve(scores.class0 = pred_ew_hsa %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_hsa %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

prolif_pr = sapply(betas, function(beta) {
  pred_ew_hsa = pred_ew_hsa %>% mutate(combined_score = ss_score + beta * prolif_score)
  res = pr.curve(scores.class0 = pred_ew_hsa %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_hsa %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_ew = as_tibble(cbind(betas, random_roc, prolif_roc, random_pr, prolif_pr))
df_ew = df_ew %>% tidyr::pivot_longer(-betas, names_to = "type", values_to = "AUC")

ggline(data = df_ew, x = "betas", y = "AUC", numeric.x.axis = TRUE, color = "type",
  plot_type = "l", xlab = TeX("$\\beta$"), ylab = "AUC (Area Under Curve)", 
  legend = "none", facet.by = "type", palette = my_palette,
  panel.labs = list(type = c("PR: calibrated + β x proliferative", 
    "ROC: calibrated + β x proliferative", "PR: calibrated + β x random", 
    "ROC: calibrated + β x random")),
  title = TeX("AUC sensitivity to $\\beta$ parameter (HSA, CASCADE 1.0)")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = -1, color = "red", size = 0.3, linetype = "dashed") + 
  geom_text(aes(x=-2, label="β = -1", y=0.25), colour="black", angle=90) + 
  grids()
```

```{r auc-sen-mw-hsa-cascade1, dpi=300, fig.align='center', cache=TRUE, fig.cap='AUC sensitivity (CASCADE 1.0, HSA synergy method, Model-wise results)'}
# Model-wise
weights = seq(from = 0, to = 1, by = 0.05)

random_roc_mw = sapply(weights, function(w) {
  pred_mw_hsa = pred_mw_hsa %>%
    mutate(weighted_prob = (1 - w) * pred_mw_hsa$synergy_prob_ss + w * pred_mw_hsa$synergy_prob_random)
  res = roc.curve(scores.class0 = pred_mw_hsa %>% pull(weighted_prob),
    weights.class0 = pred_mw_hsa %>% pull(observed))
  auc_value = res$auc
})

prolif_roc_mw = sapply(weights, function(w) {
  pred_mw_hsa = pred_mw_hsa %>%
    mutate(weighted_prob = (1 - w) * pred_mw_hsa$synergy_prob_ss + w * pred_mw_hsa$synergy_prob_prolif)
  res = roc.curve(scores.class0 = pred_mw_hsa %>% pull(weighted_prob),
    weights.class0 = pred_mw_hsa %>% pull(observed))
  auc_value = res$auc
})

random_pr_mw = sapply(weights, function(w) {
  pred_mw_hsa = pred_mw_hsa %>% 
    mutate(weighted_prob = (1 - w) * pred_mw_hsa$synergy_prob_ss + w * pred_mw_hsa$synergy_prob_random)
  res = pr.curve(scores.class0 = pred_mw_hsa %>% pull(weighted_prob), 
    weights.class0 = pred_mw_hsa %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

prolif_pr_mw = sapply(weights, function(w) {
  pred_mw_hsa = pred_mw_hsa %>% 
    mutate(weighted_prob = (1 - w) * pred_mw_hsa$synergy_prob_ss + w * pred_mw_hsa$synergy_prob_prolif)
  res = pr.curve(scores.class0 = pred_mw_hsa %>% pull(weighted_prob), 
    weights.class0 = pred_mw_hsa %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_mw = as_tibble(cbind(weights, random_roc_mw, prolif_roc_mw, random_pr_mw, prolif_pr_mw))
df_mw = df_mw %>% tidyr::pivot_longer(-weights, names_to = "type", values_to = "AUC")

ggline(data = df_mw, x = "weights", y = "AUC", numeric.x.axis = TRUE, color = "type",
  plot_type = "l", xlab = TeX("weight $w$"), ylab = "AUC (Area Under Curve)", 
  legend = "none", facet.by = "type", palette = my_palette,
  panel.labs = list(type = c("PR: (1-w) x prob(ss) + w x prob(prolif)", 
    "ROC: (1-w) x prob(ss) + w x prob(prolif)", "PR: (1-w) x prob(ss) + w x prob(random)", 
    "ROC: (1-w) x prob(ss) + w x prob(random)")), title.position = "center",
  title = TeX("AUC sensitivity to weighted average score (HSA, CASCADE 1.0)")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  grids()
```

:::{.green-box}
- There are $\beta$ values that can boost the predictive performance of the calibrated models (ensemble-wise) but no $w$ weight in the model-wise case.
- $\beta=-1$ seems to be a common value that maximizes both the ROC-AUC and the PR-AUC.
- The PR-AUC is **more sensitive** than the ROC-AUC, so a better indicator of performance.
:::

## Bliss results {-}

:::{.note}
- *Bliss* refers to the synergy method used in `Drabme` to assess the synergies from the `gitsbe` models
- We test performance using ROC and PR AUC for both the *ensemble-wise* and *model-wise* synergies from `Drabme`
- **Calibrated** models: fitted to steady state ($50$ simulations)
- **Proliferative** models: fitted to [proliferation profile](https://druglogics.github.io/druglogics-doc/training-data.html#unperturbed-condition---globaloutput-response) ($50$ simulations)
- **Random** models: produced via `abmlog` (see [here](#random-model-results) and used in `Drabme` with `synergy_method: bliss`
- `Gitsbe` models have mutations on **link operator** only
:::

Load results:
```{r Load CASCADE 1.0 Bliss results}
# 'ss' => calibrated models, 'random' => random models, 'prolif' => proliferative models

## Bliss results
ss_bliss_ensemblewise_file = paste0("results/link-only/bliss/cascade_1.0_ss_50sim_fixpoints_ensemblewise_synergies.tab")
ss_bliss_modelwise_file = paste0("results/link-only/bliss/cascade_1.0_ss_50sim_fixpoints_modelwise_synergies.tab")
prolif_bliss_ensemblewise_file = paste0("results/link-only/bliss/cascade_1.0_rand_50sim_fixpoints_ensemblewise_synergies.tab")
prolif_bliss_modelwise_file = paste0("results/link-only/bliss/cascade_1.0_rand_50sim_fixpoints_modelwise_synergies.tab")
random_bliss_ensemblewise_file = paste0("results/link-only/bliss/cascade_1.0_random_bliss_ensemblewise_synergies.tab")
random_bliss_modelwise_file = paste0("results/link-only/bliss/cascade_1.0_random_bliss_modelwise_synergies.tab")

ss_bliss_ensemblewise_synergies = emba::get_synergy_scores(ss_bliss_ensemblewise_file)
ss_bliss_modelwise_synergies = emba::get_synergy_scores(ss_bliss_modelwise_file, file_type = "modelwise")
prolif_bliss_ensemblewise_synergies = emba::get_synergy_scores(prolif_bliss_ensemblewise_file)
prolif_bliss_modelwise_synergies = emba::get_synergy_scores(prolif_bliss_modelwise_file, file_type = "modelwise")
random_bliss_ensemblewise_synergies = emba::get_synergy_scores(random_bliss_ensemblewise_file)
random_bliss_modelwise_synergies = emba::get_synergy_scores(random_bliss_modelwise_file, file_type = "modelwise")

# calculate probability of synergy in the modelwise results
ss_bliss_modelwise_synergies = ss_bliss_modelwise_synergies %>% 
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
prolif_bliss_modelwise_synergies = prolif_bliss_modelwise_synergies %>%
  mutate(synergy_prob_prolif = synergies/(synergies + `non-synergies`))
random_bliss_modelwise_synergies = random_bliss_modelwise_synergies %>%
  mutate(synergy_prob_random = synergies/(synergies + `non-synergies`))

# 'ew' => ensemble-wise, 'mw' => model-wise
pred_ew_bliss = bind_cols(ss_bliss_ensemblewise_synergies %>% rename(ss_score = score), 
  prolif_bliss_ensemblewise_synergies %>% select(score) %>% rename(prolif_score = score),
  random_bliss_ensemblewise_synergies %>% select(score) %>% rename(random_score = score), 
  as_tibble_col(observed, column_name = "observed"))

pred_mw_bliss = bind_cols(
  ss_bliss_modelwise_synergies %>% select(perturbation, synergy_prob_ss),
  prolif_bliss_modelwise_synergies %>% select(synergy_prob_prolif),
  random_bliss_modelwise_synergies %>% select(synergy_prob_random),
  as_tibble_col(observed, column_name = "observed"))
```

### ROC curves {-}

```{r roc-bliss-cascade1, fig.width=5, fig.height=5, dpi=300, cache=TRUE, fig.show='hold', out.width='50%', fig.cap='ROC curves (CASCADE 1.0, Bliss synergy method)', fig.align='center'}
res_ss_ew = get_roc_stats(df = pred_ew_bliss, pred_col = "ss_score", label_col = "observed")
res_prolif_ew = get_roc_stats(df = pred_ew_bliss, pred_col = "prolif_score", label_col = "observed")
res_random_ew = get_roc_stats(df = pred_ew_bliss, pred_col = "random_score", label_col = "observed")

res_ss_mw = get_roc_stats(df = pred_mw_bliss, pred_col = "synergy_prob_ss", label_col = "observed", direction = ">")
res_prolif_mw = get_roc_stats(df = pred_mw_bliss, pred_col = "synergy_prob_prolif", label_col = "observed", direction = ">")
res_random_mw = get_roc_stats(df = pred_mw_bliss, pred_col = "synergy_prob_random", label_col = "observed", direction = ">")

# Plot ROCs
plot(x = res_ss_ew$roc_stats$FPR, y = res_ss_ew$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Ensemble-wise synergies (Bliss)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = res_prolif_ew$roc_stats$FPR, y = res_prolif_ew$roc_stats$TPR, 
  lwd = 3, col = my_palette[2])
lines(x = res_random_ew$roc_stats$FPR, y = res_random_ew$roc_stats$TPR, 
  lwd = 3, col = my_palette[3])
legend('bottomright', title = 'AUC', col = my_palette[1:3], pch = 19,
  legend = c(paste(round(res_ss_ew$AUC, digits = 3), "Calibrated"), 
    paste(round(res_prolif_ew$AUC, digits = 3), "Proliferative"),
    paste(round(res_random_ew$AUC, digits = 3), "Random")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)

plot(x = res_ss_mw$roc_stats$FPR, y = res_ss_mw$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Model-wise synergies (Bliss)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = res_prolif_mw$roc_stats$FPR, y = res_prolif_mw$roc_stats$TPR, 
  lwd = 3, col = my_palette[2])
lines(x = res_random_mw$roc_stats$FPR, y = res_random_mw$roc_stats$TPR, 
  lwd = 3, col = my_palette[3])
legend('bottomright', title = 'AUC', col = my_palette[1:3], pch = 19,
  legend = c(paste(round(res_ss_mw$AUC, digits = 3), "Calibrated"),
    paste(round(res_prolif_mw$AUC, digits = 3), "Proliferative"), 
    paste(round(res_random_mw$AUC, digits = 3), "Random")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)
```

### PR curves {-}

```{r pr-bliss-cascade1, fig.width=5, fig.height=5, dpi=300, cache=TRUE, fig.show='hold', out.width='50%', fig.cap='PR curves (CASCADE 1.0, Bliss synergy method)', fig.align='center'}
pr_ss_ew_bliss = pr.curve(scores.class0 = pred_ew_bliss %>% pull(ss_score) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_bliss %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_prolif_ew_bliss = pr.curve(scores.class0 = pred_ew_bliss %>% pull(prolif_score) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_bliss %>% pull(observed), curve = TRUE)
pr_random_ew_bliss = pr.curve(scores.class0 = pred_ew_bliss %>% pull(random_score) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_bliss %>% pull(observed), curve = TRUE)

pr_ss_mw_bliss = pr.curve(scores.class0 = pred_mw_bliss %>% pull(synergy_prob_ss), 
  weights.class0 = pred_mw_bliss %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_prolif_mw_bliss = pr.curve(scores.class0 = pred_mw_bliss %>% pull(synergy_prob_prolif), 
  weights.class0 = pred_mw_bliss %>% pull(observed), curve = TRUE)
pr_random_mw_bliss = pr.curve(scores.class0 = pred_mw_bliss %>% pull(synergy_prob_random), 
  weights.class0 = pred_mw_bliss %>% pull(observed), curve = TRUE)

plot(pr_ss_ew_bliss, main = 'PR curve, Ensemble-wise synergies (Bliss)', 
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_prolif_ew_bliss, add = TRUE, color = my_palette[2])
plot(pr_random_ew_bliss, add = TRUE, color = my_palette[3])
legend('bottomright', title = 'AUC', col = my_palette[1:3], pch = 19, cex = 0.65,
  legend = c(paste(round(pr_ss_ew_bliss$auc.davis.goadrich, digits = 3), "Calibrated"), 
    paste(round(pr_prolif_ew_bliss$auc.davis.goadrich, digits = 3), "Proliferative"),
    paste(round(pr_random_ew_bliss$auc.davis.goadrich, digits = 3), "Random")))
grid(lwd = 0.5)

plot(pr_ss_mw_bliss, main = 'PR curve, Model-wise synergies (Bliss)', 
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_prolif_mw_bliss, add = TRUE, color = my_palette[2])
plot(pr_random_mw_bliss, add = TRUE, color = my_palette[3])
legend('left', title = 'AUC', col = my_palette[1:3], pch = 19, cex = 0.8,
  legend = c(paste(round(pr_ss_mw_bliss$auc.davis.goadrich, digits = 3), "Calibrated"),
    paste(round(pr_prolif_mw_bliss$auc.davis.goadrich, digits = 3), "Proliferative"),
    paste(round(pr_random_mw_bliss$auc.davis.goadrich, digits = 3), "Random")))
grid(lwd = 0.5)
```

:::{.green-box}
- Calibrated models perform a lot better than either the random or proliferative ones
:::

### AUC sensitivity {-}

Investigate same thing as described in [here](#auc-sensitivity).

```{r auc-sen-ew-bliss-cascade1, dpi=300, fig.align='center', cache=TRUE, fig.cap='AUC sensitivity (CASCADE 1.0, Bliss synergy method, Ensemble-wise results)'}
# Ensemble-wise
betas = seq(from = -20, to = 20, by = 0.1)

random_roc = sapply(betas, function(beta) {
  pred_ew_bliss = pred_ew_bliss %>% mutate(combined_score = ss_score + beta * random_score)
  res = roc.curve(scores.class0 = pred_ew_bliss %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_bliss %>% pull(observed))
  auc_value = res$auc
})

prolif_roc = sapply(betas, function(beta) {
  pred_ew_bliss = pred_ew_bliss %>% mutate(combined_score = ss_score + beta * prolif_score)
  res = roc.curve(scores.class0 = pred_ew_bliss %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_bliss %>% pull(observed))
  auc_value = res$auc
})

random_pr = sapply(betas, function(beta) {
  pred_ew_bliss = pred_ew_bliss %>% mutate(combined_score = ss_score + beta * random_score)
  res = pr.curve(scores.class0 = pred_ew_bliss %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_bliss %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

prolif_pr = sapply(betas, function(beta) {
  pred_ew_bliss = pred_ew_bliss %>% mutate(combined_score = ss_score + beta * prolif_score)
  res = pr.curve(scores.class0 = pred_ew_bliss %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_bliss %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_ew = as_tibble(cbind(betas, random_roc, prolif_roc, random_pr, prolif_pr))
df_ew = df_ew %>% tidyr::pivot_longer(-betas, names_to = "type", values_to = "AUC")

ggline(data = df_ew, x = "betas", y = "AUC", numeric.x.axis = TRUE, color = "type",
  plot_type = "l", xlab = TeX("$\\beta$"), ylab = "AUC (Area Under Curve)", 
  legend = "none", facet.by = "type", palette = my_palette,
  panel.labs = list(type = c("PR: calibrated + β x proliferative", 
    "ROC: calibrated + β x proliferative", "PR: calibrated + β x random", 
    "ROC: calibrated + β x random")),
  title = TeX("AUC sensitivity to $\\beta$ parameter (Bliss, CASCADE 1.0)")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = -1, color = "red", size = 0.3, linetype = "dashed") + 
  geom_text(aes(x=-2, label="β = -1", y=0.25), colour="black", angle=90) + 
  grids()
```

```{r auc-sen-mw-bliss-cascade1, dpi=300, fig.align='center', cache=TRUE, fig.cap='AUC sensitivity (CASCADE 1.0, Bliss synergy method, Model-wise results)'}
# Model-wise
weights = seq(from = 0, to = 1, by = 0.05)

random_roc_mw = sapply(weights, function(w) {
  pred_mw_bliss = pred_mw_bliss %>%
    mutate(weighted_prob = (1 - w) * pred_mw_bliss$synergy_prob_ss + w * pred_mw_bliss$synergy_prob_random)
  res = roc.curve(scores.class0 = pred_mw_bliss %>% pull(weighted_prob),
    weights.class0 = pred_mw_bliss %>% pull(observed))
  auc_value = res$auc
})

prolif_roc_mw = sapply(weights, function(w) {
  pred_mw_bliss = pred_mw_bliss %>%
    mutate(weighted_prob = (1 - w) * pred_mw_bliss$synergy_prob_ss + w * pred_mw_bliss$synergy_prob_prolif)
  res = roc.curve(scores.class0 = pred_mw_bliss %>% pull(weighted_prob),
    weights.class0 = pred_mw_bliss %>% pull(observed))
  auc_value = res$auc
})

random_pr_mw = sapply(weights, function(w) {
  pred_mw_bliss = pred_mw_bliss %>% 
    mutate(weighted_prob = (1 - w) * pred_mw_bliss$synergy_prob_ss + w * pred_mw_bliss$synergy_prob_random)
  res = pr.curve(scores.class0 = pred_mw_bliss %>% pull(weighted_prob), 
    weights.class0 = pred_mw_bliss %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

prolif_pr_mw = sapply(weights, function(w) {
  pred_mw_bliss = pred_mw_bliss %>% 
    mutate(weighted_prob = (1 - w) * pred_mw_bliss$synergy_prob_ss + w * pred_mw_bliss$synergy_prob_prolif)
  res = pr.curve(scores.class0 = pred_mw_bliss %>% pull(weighted_prob), 
    weights.class0 = pred_mw_bliss %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_mw = as_tibble(cbind(weights, random_roc_mw, prolif_roc_mw, random_pr_mw, prolif_pr_mw))
df_mw = df_mw %>% tidyr::pivot_longer(-weights, names_to = "type", values_to = "AUC")

ggline(data = df_mw, x = "weights", y = "AUC", numeric.x.axis = TRUE, color = "type",
  plot_type = "l", xlab = TeX("weight $w$"), ylab = "AUC (Area Under Curve)", 
  legend = "none", facet.by = "type", palette = my_palette,
  panel.labs = list(type = c("PR: (1-w) x prob(ss) + w x prob(prolif)", 
    "ROC: (1-w) x prob(ss) + w x prob(prolif)", "PR: (1-w) x prob(ss) + w x prob(random)", 
    "ROC: (1-w) x prob(ss) + w x prob(random)")), title.position = "center",
  title = TeX("AUC sensitivity to weighted average score (Bliss, CASCADE 1.0)")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  grids()
```

:::{.green-box}
- There are $\beta$ values that can boost the predictive performance of the calibrated models (ensemble-wise) but no $w$ weight in the model-wise case.
- $\beta=-1$ seems to be a common value that maximizes both the ROC-AUC and the PR-AUC.
- The PR-AUC is **more sensitive** than the ROC-AUC, so a better indicator of performance.
:::

## Correlation {-}

We test for correlation between all the synergy predictor results shown in the previous curves.
This means *ensemble-wise* vs *model-wise*, *random* models vs *proliferative* vs *calibrated (ss)* models and *HSA* vs *Bliss* synergy assessment.
*P-values* are represented at 3 significant levels: $0.05, 0.01, 0.001$ (\*, \*\*, \*\*\*) and the correlation coefficient is calculated using Kendall's *tau* statistic.

```{r cor-plot-cascade1, dpi=300, warning=FALSE, cache=TRUE, fig.align='center', fig.cap='Correlation Plot for CASCADE 1.0 Results'}
synergy_scores = bind_cols(
  pred_ew_hsa %>% select(ss_score, prolif_score, random_score) %>% rename(ss_ew_hsa = ss_score, prolif_ew_hsa = prolif_score, random_ew_hsa = random_score),
  pred_ew_bliss %>% select(ss_score, prolif_score, random_score) %>% rename(ss_ew_bliss = ss_score, prolif_ew_bliss = prolif_score, random_ew_bliss = random_score),
  pred_mw_hsa %>% select(synergy_prob_ss, synergy_prob_prolif, synergy_prob_random) %>% 
    rename(ss_mw_hsa = synergy_prob_ss, prolif_mw_hsa = synergy_prob_prolif, random_mw_hsa = synergy_prob_random),
  pred_mw_bliss %>% select(synergy_prob_ss, synergy_prob_prolif, synergy_prob_random) %>% 
    rename(ss_mw_bliss = synergy_prob_ss, prolif_mw_bliss = synergy_prob_prolif, random_mw_bliss = synergy_prob_random)
  )

M = cor(synergy_scores, method = "kendall")
res = cor.mtest(synergy_scores, method = "kendall")
corrplot(corr = M, type = "upper", p.mat = res$p, sig.level = c(.001, .01, .05), 
  pch.cex = 1, pch.col = "white", insig = "label_sig", tl.col = "black", tl.srt = 45)
```

:::{.green-box}
- **Model-wise** don't correlate at all with **ensemble-wise** results (*topright* part of the correlation plot).
- **HSA and Bliss results correlate**, higher for the model-wise (*bottomright*) than the ensemble-wise results (*topleft*).
- **Calibrated** results don't correlate with either **random or proliferative** results in most cases, while there is **stronger correlation between the proliferative and random models results**.
:::

## Fitness Evolution {-}

Results are from the simulation with $50$ Gitsbe simulations, fitting to steady state (**calibrated models**) and *HSA* Drabme synergy assessment.
We show only $10$ simulations - the first ones that spanned the maximum defined generations in the configuration ($20$), meaning that they did not surpass the target fitness threhold specified ($0.99$).
Each data point is the average fitness in that generation out of $20$ models.

```{r fit-evolution, dpi=300, fig.align='center', cache=TRUE, fig.cap='Fitness Evolution (10 simulations, CASCADE 1.0, HSA synergy method)'}
read_summary_file = function(file_name) {
  lines = readr::read_lines(file = file_name, skip = 5, skip_empty_rows = TRUE)
  
  data_list = list()
  index = 1
  
  gen_fit_list = list()
  gen_index = 1
  for (line_index in 1:length(lines)) {
    line = lines[line_index]
    if (stringr::str_detect(string = line, pattern = "Simulation")) {
      data_list[[index]] = bind_cols(gen_fit_list)
      index = index + 1
      
      gen_fit_list = list()
      gen_index = 1
    } else { # read fitness values
      gen_fit_list[[gen_index]] = as_tibble_col(as.numeric(unlist(strsplit(line, split = '\t'))))
      gen_index = gen_index + 1
    }
  }
  
  # add the last simulation's values
  data_list[[index]] = bind_cols(gen_fit_list)
  
  return(data_list)
}

fitness_summary_file = paste0("results/link-only/hsa/cascade_1.0_ss_50sim_fixpoints_summary.txt")
fit_res = read_summary_file(file_name = fitness_summary_file)

first_sim_data = colMeans(fit_res[[1]])
plot(1:length(first_sim_data), y = first_sim_data, ylim = c(0,1), 
  xlim = c(0,20), type = 'l', lwd = 1.5, 
  main = 'Fitness Evolution across Generations', xlab = 'Generations',
  ylab = 'Average Fitness', col = usefun:::colors.100[1])
index = 2
for (fit_data in fit_res) {
  if (index > 10) break
  if (ncol(fit_data) != 20) next
  mean_fit_per_gen = colMeans(fit_data)
  lines(x = 1:length(mean_fit_per_gen), y = mean_fit_per_gen, lwd = 1.5,
    col = usefun:::colors.100[index])
  index = index + 1
}
grid(lwd = 0.5)
```

If we plot the average fitness from all $50$ simulations (per generation) and add the standard deviation errorbars we have:

```{r fit-evolution-2, dpi=300, fig.align='center', warning=FALSE, cache=TRUE, fig.cap='Fitness Evolution (50 simulations, CASCADE 1.0, HSA synergy method)'}
# rows = simulations, columns = generations
# value in (sim,gen) cell = average fitness of models in that particular (sim,gen) combination
avg_fit = do.call(dplyr::bind_rows, sapply(fit_res, colMeans))
colnames(avg_fit) = 1:ncol(avg_fit)

avg_fit_long = avg_fit %>% pivot_longer(cols = everything()) %>% mutate(name = as.integer(name))

ggline(data = avg_fit_long, x = "name", y = "value", color = my_palette[2],
  add = "mean_sd", add.params = list(color = "black"), 
  main = "Fitness Evolution across Generations", 
  xlab = "Generations", ylab = "Fitness") + 
  theme(plot.title = element_text(hjust = 0.5)) + grids()

# DIY way:
# df = avg_fit_long %>% group_by(name) %>%
#   summarise(median = median(value, na.rm = TRUE),
#     mean = mean(value, na.rm = TRUE),
#     sd = sd(value, na.rm = TRUE))
# 
# ggplot(data = df, aes(x=name, y=mean)) +
#   ggtitle("Fitness Evolution across Generations") +
#   xlab("Generations") + ylab("Fitness") +
#   geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2) +
#   geom_line(color='steelblue') +
#   geom_point(color='steelblue') +
#   theme_pubr() + theme(plot.title = element_text(hjust = 0.5)) +
#   grids()
```

# CASCADE 2.0 Analysis (Link Operator Mutations) {-}

:::{.blue-box}
Performance of automatically parameterized models against SINTEF dataset [@Flobak2019]
:::

## HSA results {-}

:::{.note}
- *HSA* refers to the synergy method used in `Drabme` to assess the synergies from the `gitsbe` models
- We test performance using ROC and PR AUC for both the *ensemble-wise* and *model-wise* synergies from `Drabme`
- **Calibrated** models: fitted to steady state ($50,100,150,200$ simulations)
- **Proliferative** models: fitted to [proliferation profile](https://druglogics.github.io/druglogics-doc/training-data.html#unperturbed-condition---globaloutput-response) ($150$ simulations)
- **Random** models: produced via `abmlog` (see [here](#random-model-results) and used in `Drabme` with `synergy_method: hsa`
- `Gitsbe` models have mutations on **link operator** only
:::

Load results:

```{r load CASCADE 2.0 HSA results}
# 'ss' => calibrated models, 'random' => random models, 'prolif' => proliferative models
# 'ew' => ensemble-wise, 'mw' => model-wise

## HSA results
ss_hsa_ensemblewise_50sim_file = paste0("results/link-only/hsa/cascade_2.0_ss_50sim_fixpoints_ensemblewise_synergies.tab")
ss_hsa_modelwise_50sim_file = paste0("results/link-only/hsa/cascade_2.0_ss_50sim_fixpoints_modelwise_synergies.tab")
ss_hsa_ensemblewise_100sim_file = paste0("results/link-only/hsa/cascade_2.0_ss_100sim_fixpoints_ensemblewise_synergies.tab")
ss_hsa_modelwise_100sim_file = paste0("results/link-only/hsa/cascade_2.0_ss_100sim_fixpoints_modelwise_synergies.tab")
ss_hsa_ensemblewise_150sim_file = paste0("results/link-only/hsa/cascade_2.0_ss_150sim_fixpoints_ensemblewise_synergies.tab")
ss_hsa_modelwise_150sim_file = paste0("results/link-only/hsa/cascade_2.0_ss_150sim_fixpoints_modelwise_synergies.tab")
ss_hsa_ensemblewise_200sim_file = paste0("results/link-only/hsa/cascade_2.0_ss_200sim_fixpoints_ensemblewise_synergies.tab")
ss_hsa_modelwise_200sim_file = paste0("results/link-only/hsa/cascade_2.0_ss_200sim_fixpoints_modelwise_synergies.tab")
prolif_hsa_ensemblewise_file = paste0("results/link-only/hsa/cascade_2.0_rand_150sim_fixpoints_hsa_ensemblewise_synergies.tab")
prolif_hsa_modelwise_file = paste0("results/link-only/hsa/cascade_2.0_rand_150sim_fixpoints_hsa_modelwise_synergies.tab")
random_hsa_ensemblewise_file = paste0("results/link-only/hsa/cascade_2.0_random_ensemblewise_synergies.tab")
random_hsa_modelwise_file = paste0("results/link-only/hsa/cascade_2.0_random_modelwise_synergies.tab")

ss_hsa_ensemblewise_synergies_50sim = emba::get_synergy_scores(ss_hsa_ensemblewise_50sim_file)
ss_hsa_modelwise_synergies_50sim = emba::get_synergy_scores(ss_hsa_modelwise_50sim_file, file_type = "modelwise")
ss_hsa_ensemblewise_synergies_100sim = emba::get_synergy_scores(ss_hsa_ensemblewise_100sim_file)
ss_hsa_modelwise_synergies_100sim = emba::get_synergy_scores(ss_hsa_modelwise_100sim_file, file_type = "modelwise")
ss_hsa_ensemblewise_synergies_150sim = emba::get_synergy_scores(ss_hsa_ensemblewise_150sim_file)
ss_hsa_modelwise_synergies_150sim = emba::get_synergy_scores(ss_hsa_modelwise_150sim_file, file_type = "modelwise")
ss_hsa_ensemblewise_synergies_200sim = emba::get_synergy_scores(ss_hsa_ensemblewise_200sim_file)
ss_hsa_modelwise_synergies_200sim = emba::get_synergy_scores(ss_hsa_modelwise_200sim_file, file_type = "modelwise")
prolif_hsa_ensemblewise_synergies_150sim = emba::get_synergy_scores(prolif_hsa_ensemblewise_file)
prolif_hsa_modelwise_synergies_150sim = emba::get_synergy_scores(prolif_hsa_modelwise_file, file_type = "modelwise")
random_hsa_ensemblewise_synergies = emba::get_synergy_scores(random_hsa_ensemblewise_file)
random_hsa_modelwise_synergies = emba::get_synergy_scores(random_hsa_modelwise_file, file_type = "modelwise")

# calculate probability of synergy in the modelwise results
ss_hsa_modelwise_synergies_50sim = ss_hsa_modelwise_synergies_50sim %>% 
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
ss_hsa_modelwise_synergies_100sim = ss_hsa_modelwise_synergies_100sim %>% 
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
ss_hsa_modelwise_synergies_150sim = ss_hsa_modelwise_synergies_150sim %>% 
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
ss_hsa_modelwise_synergies_200sim = ss_hsa_modelwise_synergies_200sim %>% 
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
prolif_hsa_modelwise_synergies_150sim = prolif_hsa_modelwise_synergies_150sim %>%
  mutate(synergy_prob_prolif = synergies/(synergies + `non-synergies`))
random_hsa_modelwise_synergies = random_hsa_modelwise_synergies %>%
  mutate(synergy_prob_random = synergies/(synergies + `non-synergies`))

observed_synergies_file = paste0("results/observed_synergies_cascade_2.0")
observed_synergies = get_observed_synergies(observed_synergies_file)
# 1 (positive/observed synergy) or 0 (negative/not observed) for all tested drug combinations
observed = sapply(random_hsa_modelwise_synergies$perturbation %in% observed_synergies, as.integer)

# 'ew' => ensemble-wise, 'mw' => model-wise
pred_ew_hsa = bind_cols(
  ss_hsa_ensemblewise_synergies_50sim %>% select(score) %>% rename(ss_score_50sim = score),
  ss_hsa_ensemblewise_synergies_100sim %>% select(score) %>% rename(ss_score_100sim = score),
  ss_hsa_ensemblewise_synergies_150sim %>% select(score) %>% rename(ss_score_150sim = score),
  ss_hsa_ensemblewise_synergies_200sim %>% select(score) %>% rename(ss_score_200sim = score),
  prolif_hsa_ensemblewise_synergies_150sim %>% select(score) %>% rename(prolif_score_150sim = score),
  random_hsa_ensemblewise_synergies %>% select(score) %>% rename(random_score = score), 
  as_tibble_col(observed, column_name = "observed"))

pred_mw_hsa = bind_cols(
  ss_hsa_modelwise_synergies_50sim %>% select(perturbation, synergy_prob_ss) %>% rename(synergy_prob_ss_50sim = synergy_prob_ss),
  ss_hsa_modelwise_synergies_100sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_ss_100sim = synergy_prob_ss),
  ss_hsa_modelwise_synergies_150sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_ss_150sim = synergy_prob_ss),
  ss_hsa_modelwise_synergies_200sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_ss_200sim = synergy_prob_ss),
  prolif_hsa_modelwise_synergies_150sim %>% select(synergy_prob_prolif) %>% rename(synergy_prob_prolif_150sim = synergy_prob_prolif),
  random_hsa_modelwise_synergies %>% select(synergy_prob_random),
  as_tibble_col(observed, column_name = "observed"))
```

### ROC curves {-}

```{r roc-hsa-cascade2, fig.width=5, fig.height=5, dpi=300, cache=TRUE, fig.show='hold', out.width='50%', fig.cap='ROC curves (CASCADE 2.0, HSA synergy method)', fig.align='center'}
res_ss_ew_50sim = get_roc_stats(df = pred_ew_hsa, pred_col = "ss_score_50sim", label_col = "observed")
res_ss_ew_100sim = get_roc_stats(df = pred_ew_hsa, pred_col = "ss_score_100sim", label_col = "observed")
res_ss_ew_150sim = get_roc_stats(df = pred_ew_hsa, pred_col = "ss_score_150sim", label_col = "observed")
res_ss_ew_200sim = get_roc_stats(df = pred_ew_hsa, pred_col = "ss_score_200sim", label_col = "observed")
res_prolif_ew_150sim = get_roc_stats(df = pred_ew_hsa, pred_col = "prolif_score_150sim", label_col = "observed")
res_random_ew = get_roc_stats(df = pred_ew_hsa, pred_col = "random_score", label_col = "observed")

res_ss_mw_50sim = get_roc_stats(df = pred_mw_hsa, pred_col = "synergy_prob_ss_50sim", label_col = "observed", direction = ">")
res_ss_mw_100sim = get_roc_stats(df = pred_mw_hsa, pred_col = "synergy_prob_ss_100sim", label_col = "observed", direction = ">")
res_ss_mw_150sim = get_roc_stats(df = pred_mw_hsa, pred_col = "synergy_prob_ss_150sim", label_col = "observed", direction = ">")
res_ss_mw_200sim = get_roc_stats(df = pred_mw_hsa, pred_col = "synergy_prob_ss_200sim", label_col = "observed", direction = ">")
res_prolif_mw_150sim = get_roc_stats(df = pred_mw_hsa, pred_col = "synergy_prob_prolif_150sim", label_col = "observed", direction = ">")
res_random_mw = get_roc_stats(df = pred_mw_hsa, pred_col = "synergy_prob_random", label_col = "observed", direction = ">")

# Plot ROCs
plot(x = res_ss_ew_50sim$roc_stats$FPR, y = res_ss_ew_50sim$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Ensemble-wise synergies (HSA)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = res_ss_ew_100sim$roc_stats$FPR, y = res_ss_ew_100sim$roc_stats$TPR, 
  lwd = 3, col = my_palette[2])
lines(x = res_ss_ew_150sim$roc_stats$FPR, y = res_ss_ew_150sim$roc_stats$TPR, 
  lwd = 3, col = my_palette[3])
lines(x = res_ss_ew_200sim$roc_stats$FPR, y = res_ss_ew_200sim$roc_stats$TPR, 
  lwd = 3, col = my_palette[4])
lines(x = res_prolif_ew_150sim$roc_stats$FPR, y = res_prolif_ew_150sim$roc_stats$TPR, 
  lwd = 3, col = my_palette[5])
lines(x = res_random_ew$roc_stats$FPR, y = res_random_ew$roc_stats$TPR, 
  lwd = 3, col = my_palette[6])
legend('bottomright', title = 'AUC', col = my_palette[1:6], pch = 19,
  legend = c(paste(round(res_ss_ew_50sim$AUC, digits = 3), "Calibrated (50 sim)"),
    paste(round(res_ss_ew_100sim$AUC, digits = 3), "Calibrated (100 sim)"),
    paste(round(res_ss_ew_150sim$AUC, digits = 3), "Calibrated (150 sim)"),
    paste(round(res_ss_ew_200sim$AUC, digits = 3), "Calibrated (200 sim)"),
    paste(round(res_prolif_ew_150sim$AUC, digits = 3), "Proliferative (150 sim)"),
    paste(round(res_random_ew$AUC, digits = 3), "Random")), cex = 0.9)
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)

plot(x = res_ss_mw_50sim$roc_stats$FPR, y = res_ss_mw_50sim$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Model-wise synergies (HSA)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = res_ss_mw_100sim$roc_stats$FPR, y = res_ss_mw_100sim$roc_stats$TPR, 
  lwd = 3, col = my_palette[2])
lines(x = res_ss_mw_150sim$roc_stats$FPR, y = res_ss_mw_150sim$roc_stats$TPR, 
  lwd = 3, col = my_palette[3])
lines(x = res_ss_mw_200sim$roc_stats$FPR, y = res_ss_mw_200sim$roc_stats$TPR, 
  lwd = 3, col = my_palette[4])
lines(x = res_prolif_mw_150sim$roc_stats$FPR, y = res_prolif_mw_150sim$roc_stats$TPR, 
  lwd = 3, col = my_palette[5])
lines(x = res_random_mw$roc_stats$FPR, y = res_random_mw$roc_stats$TPR, 
  lwd = 3, col = my_palette[6])
legend('bottomright', title = 'AUC', col = my_palette[1:6], pch = 19,
  legend = c(paste(round(res_ss_mw_50sim$AUC, digits = 3), "Calibrated (50 sim)"),
    paste(round(res_ss_mw_100sim$AUC, digits = 3), "Calibrated (100 sim)"),
    paste(round(res_ss_mw_150sim$AUC, digits = 3), "Calibrated (150 sim)"),
    paste(round(res_ss_mw_200sim$AUC, digits = 3), "Calibrated (200 sim)"),
    paste(round(res_prolif_mw_150sim$AUC, digits = 3), "Proliferative (150 sim)"),
    paste(round(res_random_mw$AUC, digits = 3), "Random")), cex = 0.9)
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)
```

### PR curves {-}

```{r pr-hsa-cascade2, fig.width=5, fig.height=5, dpi=300, cache=TRUE, fig.show='hold', out.width='50%', fig.cap='PR curves (CASCADE 2.0, HSA synergy method)', fig.align='center'}
pr_ss_ew_hsa_50sim = pr.curve(scores.class0 = pred_ew_hsa %>% pull(ss_score_50sim) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_hsa %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_ss_ew_hsa_100sim = pr.curve(scores.class0 = pred_ew_hsa %>% pull(ss_score_100sim) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_hsa %>% pull(observed), curve = TRUE)
pr_ss_ew_hsa_150sim = pr.curve(scores.class0 = pred_ew_hsa %>% pull(ss_score_150sim) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_hsa %>% pull(observed), curve = TRUE)
pr_ss_ew_hsa_200sim = pr.curve(scores.class0 = pred_ew_hsa %>% pull(ss_score_200sim) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_hsa %>% pull(observed), curve = TRUE)
pr_prolif_ew_hsa_150sim = pr.curve(scores.class0 = pred_ew_hsa %>% pull(prolif_score_150sim) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_hsa %>% pull(observed), curve = TRUE)
pr_random_ew_hsa = pr.curve(scores.class0 = pred_ew_hsa %>% pull(random_score) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_hsa %>% pull(observed), curve = TRUE)

pr_ss_mw_hsa_50sim = pr.curve(scores.class0 = pred_mw_hsa %>% pull(synergy_prob_ss_50sim), 
  weights.class0 = pred_mw_hsa %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_ss_mw_hsa_100sim = pr.curve(scores.class0 = pred_mw_hsa %>% pull(synergy_prob_ss_100sim), 
  weights.class0 = pred_mw_hsa %>% pull(observed), curve = TRUE)
pr_ss_mw_hsa_150sim = pr.curve(scores.class0 = pred_mw_hsa %>% pull(synergy_prob_ss_150sim), 
  weights.class0 = pred_mw_hsa %>% pull(observed), curve = TRUE)
pr_ss_mw_hsa_200sim = pr.curve(scores.class0 = pred_mw_hsa %>% pull(synergy_prob_ss_200sim), 
  weights.class0 = pred_mw_hsa %>% pull(observed), curve = TRUE)
pr_prolif_mw_hsa_150sim = pr.curve(scores.class0 = pred_mw_hsa %>% pull(synergy_prob_prolif_150sim), 
  weights.class0 = pred_mw_hsa %>% pull(observed), curve = TRUE)
pr_random_mw_hsa = pr.curve(scores.class0 = pred_mw_hsa %>% pull(synergy_prob_random), 
  weights.class0 = pred_mw_hsa %>% pull(observed), curve = TRUE)

plot(pr_ss_ew_hsa_50sim, main = 'PR curve, Ensemble-wise synergies (HSA)', 
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_ss_ew_hsa_100sim, add = TRUE, color = my_palette[2])
plot(pr_ss_ew_hsa_150sim, add = TRUE, color = my_palette[3])
plot(pr_ss_ew_hsa_200sim, add = TRUE, color = my_palette[4])
plot(pr_prolif_ew_hsa_150sim, add = TRUE, color = my_palette[5])
plot(pr_random_ew_hsa, add = TRUE, color = my_palette[6])
legend('topright', title = 'AUC', col = my_palette[1:6], pch = 19,
  legend = c(paste(round(pr_ss_ew_hsa_50sim$auc.davis.goadrich, digits = 3), "Calibrated (50 sim)"),
    paste(round(pr_ss_ew_hsa_100sim$auc.davis.goadrich, digits = 3), "Calibrated (100 sim)"),
    paste(round(pr_ss_ew_hsa_150sim$auc.davis.goadrich, digits = 3), "Calibrated (150 sim)"),
    paste(round(pr_ss_ew_hsa_200sim$auc.davis.goadrich, digits = 3), "Calibrated (200 sim)"),
    paste(round(pr_prolif_ew_hsa_150sim$auc.davis.goadrich, digits = 3), "Proliferative (150 sim)"),
    paste(round(pr_random_ew_hsa$auc.davis.goadrich, digits = 3), "Random")))
grid(lwd = 0.5)

plot(pr_ss_mw_hsa_50sim, main = 'PR curve, Model-wise synergies (HSA)',
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_ss_mw_hsa_100sim, add = TRUE, color = my_palette[2])
plot(pr_ss_mw_hsa_150sim, add = TRUE, color = my_palette[3])
plot(pr_ss_mw_hsa_200sim, add = TRUE, color = my_palette[4])
plot(pr_prolif_mw_hsa_150sim, add = TRUE, color = my_palette[5])
plot(pr_random_mw_hsa, add = TRUE, color = my_palette[6])
legend('topright', title = 'AUC', col = my_palette[1:6], pch = 19,
  legend = c(paste(round(pr_ss_mw_hsa_50sim$auc.davis.goadrich, digits = 3), "Calibrated (50 sim)"),
    paste(round(pr_ss_mw_hsa_100sim$auc.davis.goadrich, digits = 3), "Calibrated (100 sim)"),
    paste(round(pr_ss_mw_hsa_150sim$auc.davis.goadrich, digits = 3), "Calibrated (150 sim)"),
    paste(round(pr_ss_mw_hsa_200sim$auc.davis.goadrich, digits = 3), "Calibrated (200 sim)"),
    paste(round(pr_prolif_mw_hsa_150sim$auc.davis.goadrich, digits = 3), "Proliferative (200 sim)"),
    paste(round(pr_random_mw_hsa$auc.davis.goadrich, digits = 3), "Random")))
grid(lwd = 0.5)
```

:::{.green-box}
- To minimize the resulting performance variance, $150$ seems to be a good number of `Gitsbe` simulations to run for the CASCADE 2.0 network.
- The PR curves show that the **performance of each individual predictor is poor** compared to the baseline.
Someone looking at the ROC curves only might reach a different conclusion.
- *Proliferative* and *random* models perform almost equally well to *calibrated* models.
- The *model-wise* approach produces slightly better ROC results than the *ensemble-wise* approach
:::

### AUC sensitivity {-}

Investigate same thing as described in [here](#auc-sensitivity).
This is very crucial since the PR performance is poor for the individual predictors, but a combined predictor might be able to counter this.
We will combine the synergy scores from the *random* and *proliferative* simulations with the results from the *calibrated* Gitsbe simulations ($150$).

```{r auc-sen-ew-hsa-cascade2, dpi=300, fig.align='center', cache=TRUE, fig.cap='AUC sensitivity (CASCADE 2.0, HSA synergy method, Ensemble-wise results)'}
# Ensemble-wise
betas = seq(from = -7.5, to = 7.5, by = 0.1)

random_roc = sapply(betas, function(beta) {
  pred_ew_hsa = pred_ew_hsa %>% mutate(combined_score = ss_score_150sim + beta * random_score)
  res = roc.curve(scores.class0 = pred_ew_hsa %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_hsa %>% pull(observed))
  auc_value = res$auc
})

prolif_roc = sapply(betas, function(beta) {
  pred_ew_hsa = pred_ew_hsa %>% mutate(combined_score = ss_score_150sim + beta * prolif_score_150sim)
  res = roc.curve(scores.class0 = pred_ew_hsa %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_hsa %>% pull(observed))
  auc_value = res$auc
})

random_pr = sapply(betas, function(beta) {
  pred_ew_hsa = pred_ew_hsa %>% mutate(combined_score = ss_score_150sim + beta * random_score)
  res = pr.curve(scores.class0 = pred_ew_hsa %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_hsa %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

prolif_pr = sapply(betas, function(beta) {
  pred_ew_hsa = pred_ew_hsa %>% mutate(combined_score = ss_score_150sim + beta * prolif_score_150sim)
  res = pr.curve(scores.class0 = pred_ew_hsa %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_hsa %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_ew = as_tibble(cbind(betas, random_roc, prolif_roc, random_pr, prolif_pr))
df_ew = df_ew %>% tidyr::pivot_longer(-betas, names_to = "type", values_to = "AUC")

ggline(data = df_ew, x = "betas", y = "AUC", numeric.x.axis = TRUE, color = "type",
  plot_type = "l", xlab = TeX("$\\beta$"), ylab = "AUC (Area Under Curve)", 
  legend = "none", facet.by = "type", palette = my_palette,
  panel.labs = list(type = c("PR: calibrated + β x proliferative", 
    "ROC: calibrated + β x proliferative", "PR: calibrated + β x random", 
    "ROC: calibrated + β x random")),
  title = TeX("AUC sensitivity to $\\beta$ parameter (HSA, CASCADE 2.0)")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = -1, color = "red", size = 0.3, linetype = "dashed") + 
  geom_text(aes(x=-1.5, label="β = -1", y=0.25), colour="black", angle=90) +
  grids()
```

```{r auc-sen-mw-hsa-cascade2, dpi=300, fig.align='center', cache=TRUE, fig.cap='AUC sensitivity (CASCADE 2.0, HSA synergy method, Model-wise results)'}
# Model-wise
weights = seq(from = 0, to = 1, by = 0.05)

random_roc_mw = sapply(weights, function(w) {
  pred_mw_hsa = pred_mw_hsa %>%
    mutate(weighted_prob = (1 - w) * pred_mw_hsa$synergy_prob_ss_150sim + w * pred_mw_hsa$synergy_prob_random)
  res = roc.curve(scores.class0 = pred_mw_hsa %>% pull(weighted_prob),
    weights.class0 = pred_mw_hsa %>% pull(observed))
  auc_value = res$auc
})

prolif_roc_mw = sapply(weights, function(w) {
  pred_mw_hsa = pred_mw_hsa %>%
    mutate(weighted_prob = (1 - w) * pred_mw_hsa$synergy_prob_ss_150sim + w * pred_mw_hsa$synergy_prob_prolif_150sim)
  res = roc.curve(scores.class0 = pred_mw_hsa %>% pull(weighted_prob),
    weights.class0 = pred_mw_hsa %>% pull(observed))
  auc_value = res$auc
})

random_pr_mw = sapply(weights, function(w) {
  pred_mw_hsa = pred_mw_hsa %>% 
    mutate(weighted_prob = (1 - w) * pred_mw_hsa$synergy_prob_ss_150sim + w * pred_mw_hsa$synergy_prob_random)
  res = pr.curve(scores.class0 = pred_mw_hsa %>% pull(weighted_prob), 
    weights.class0 = pred_mw_hsa %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

prolif_pr_mw = sapply(weights, function(w) {
  pred_mw_hsa = pred_mw_hsa %>% 
    mutate(weighted_prob = (1 - w) * pred_mw_hsa$synergy_prob_ss_150sim + w * pred_mw_hsa$synergy_prob_prolif_150sim)
  res = pr.curve(scores.class0 = pred_mw_hsa %>% pull(weighted_prob), 
    weights.class0 = pred_mw_hsa %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_mw = as_tibble(cbind(weights, random_roc_mw, prolif_roc_mw, random_pr_mw, prolif_pr_mw))
df_mw = df_mw %>% tidyr::pivot_longer(-weights, names_to = "type", values_to = "AUC")

ggline(data = df_mw, x = "weights", y = "AUC", numeric.x.axis = TRUE, color = "type",
  plot_type = "l", xlab = TeX("weight $w$"), ylab = "AUC (Area Under Curve)", 
  legend = "none", facet.by = "type", palette = my_palette,
  panel.labs = list(type = c("PR: (1-w) x prob(ss) + w x prob(prolif)", 
    "ROC: (1-w) x prob(ss) + w x prob(prolif)", "PR: (1-w) x prob(ss) + w x prob(random)", 
    "ROC: (1-w) x prob(ss) + w x prob(random)")), title.position = "center",
  title = TeX("AUC sensitivity to weighted average score (HSA, CASCADE 2.0)")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  grids()
```

:::{.green-box}
- No added benefit when using the *model-wise* approach.
- Neither the random nor the proliferative models bring any significant change to the prediction performance of the calibrated models (ensemble-wise). 
- Only the proliferative models seem to add a small contribution to the calibrated models performance (top-right panel => ROC-AUC increases, PR-AUC is insignificantly changed nonetheless).
- The $\beta_{best}$ that maximizes the ROC and PR AUC for the **combination of proliferative and calibrated models** and is equal to $\beta_{best}=-0.3$.
For $\beta=-1$ we do not observe performance improvement in this case.
:::

### Logistic Regression {-}

We tried fitting a model using logistic regression as a different approach to combine/augment the results from calibrated simulations with the proliferative ones (for the HSA-assessed ensemble-wise results where there was a minimal benefit).

```{r logistic regression model, warning=FALSE}
model = glm(formula = observed ~ ss_score_150sim + prolif_score_150sim - 1, data = pred_ew_hsa, family = binomial())
model_tidy = broom::tidy(model)
coef1 = model_tidy %>% filter(term == "ss_score_150sim") %>% pull(estimate)
coef2 = model_tidy %>% filter(term == "prolif_score_150sim") %>% pull(estimate)
pred_ew_hsa = pred_ew_hsa %>% mutate(glm = coef1 * ss_score_150sim + coef2 * prolif_score_150sim)
res_roc = get_roc_stats(df = pred_ew_hsa, pred_col = "glm", label_col = "observed")
res_pr = pr.curve(scores.class0 = pred_ew_hsa %>% pull(glm) %>% (function(x) {-x}), weights.class0 = pred_ew_hsa %>% pull(observed))
```

The model with the coefficients is as follows (note that adding an intercept makes ROC AUC result considerably worse):
```{r equation, results='asis'}
extract_eq(model, use_coefs = TRUE)
```

:::{.orange-box}
The ROC AUC produced with a logistic regression model is lower than the calibrated models (with $150$ Gitsbe simulations): **`r res_roc$AUC`** (PR-AUC is also lower: **`r res_pr$auc.davis.goadrich`**).
:::

### Regularized Logistic Regression {-}

Because the coefficient values found from the above approach are large, we try a regularized logistic regression approach using the `glmnet` R package [@R-glmnet].
We cross validate the $\lambda$ parameter and try with different $\alpha \in [0,1]$ ($\alpha=0$ means Ridge regression, $\alpha=1$ means LASSO, in between means Elastic net) while either minimizing the missclassification error (`type.measure="class"`) or maximizing the ROC-AUC (`type.measure = "auc"`).
For each respective $\alpha$ we choose the $\lambda_{min}$ as the one the minimizes the average CV error.
The intercept was again excluded as it resulted in worse AUC performance.

```{r regularized logistic regression, warning=FALSE, cache=TRUE}
x = pred_ew_hsa %>% select(ss_score_150sim, prolif_score_150sim) %>% as.matrix()
y = pred_ew_hsa %>% pull(observed)

data_list = list()
index = 1
for (i in 0:10) { # from Ridge to LASSO
  a = i/10
  for (measure in c("auc", "class")) {
    set.seed(42) # for reproducibility
    cvfit = cv.glmnet(x, y, family = "binomial", type.measure = measure, intercept = FALSE, alpha = a)
    coef_mat = coef(cvfit, s = "lambda.min")
    pred_ew_hsa = pred_ew_hsa %>% mutate(glm_reg = coef_mat[1] + coef_mat[2] * ss_score_150sim + coef_mat[3] * random_score)
    res_roc = get_roc_stats(df = pred_ew_hsa, pred_col = "glm_reg", label_col = "observed")
    pr_roc = pr.curve(scores.class0 = pred_ew_hsa %>% pull(glm_reg) %>% (function(x) {-x}), 
      weights.class0 = pred_ew_hsa %>% pull(observed))
    data_list[[index]] = as_tibble_row(list(alpha = a, measure = measure, ROC_AUC = res_roc$AUC, PR_AUC = pr_roc$auc.davis.goadrich))
    index = index + 1
  }
}

data = bind_rows(data_list)

# List the best two results
data %>% arrange(desc(ROC_AUC)) %>% slice(1:4) %>% kable()
```

:::{.orange-box}
The best ROC AUC produced with a regularized logistic regression model is also lower than the one using calibrated models alone (with $150$ Gitsbe simulations).

Note that we get warnings when using `glmnet` because of the small number of observations for the positive class (observed synergies).
Resulting coefficients vary, but tend to be either all too small or **larger on the proliferative model predictor**.
:::

### MAMSE ROC {-}

Using the `MAMSE` R package [@R-MAMSE] we try another method to combine the predictor values from the calibrated and proliferative models.
The resulting ROC curve gets a little bit distored and AUC is not statistically better from the reference sample population (i.e. the calibrated `Gitsbe` models with $150$ simulations):

```{r MAMSE-test, fig.align='center', fig.cap='Combined Ensemble-wise Classifier using MAMSE ROC (CASCADE 2.0, HSA)'}
# healthy => non-synergy, diseased => synergy
healthy = list()
healthy[[1]] = pred_ew_hsa %>% filter(observed == 0) %>% pull(ss_score_150sim)
healthy[[2]] = pred_ew_hsa %>% filter(observed == 0) %>% pull(prolif_score_150sim)

diseased = list()
diseased[[1]] = pred_ew_hsa %>% filter(observed == 1) %>% pull(ss_score_150sim)
diseased[[2]] = pred_ew_hsa %>% filter(observed == 1) %>% pull(prolif_score_150sim)

plot(roc(healthy = healthy, diseased = diseased, smalldiseased=TRUE, AUC=TRUE,
  wh=NULL, wd=NULL, FPR=NULL, method="np"))
```

## Bliss results {-}

:::{.note}
- *Bliss* refers to the synergy method used in `Drabme` to assess the synergies from the `gitsbe` models
- We test performance using ROC and PR AUC for both the *ensemble-wise* and *model-wise* synergies from `Drabme`
- **Calibrated** models: fitted to steady state ($50,100,150,200$ simulations)
- **Proliferative** models: fitted to [proliferation profile](https://druglogics.github.io/druglogics-doc/training-data.html#unperturbed-condition---globaloutput-response) ($150$ simulations)
- **Random** models: produced via `abmlog` (see [here](#random-model-results) and used in `Drabme` with `synergy_method: bliss`
- `Gitsbe` models have mutations on **link operator** only
:::

Load results:
```{r Load CASCADE 2.0 Bliss results}
# 'ss' => calibrated models, 'random' => random models

## Bliss results
ss_bliss_ensemblewise_50sim_file = paste0("results/link-only/bliss/cascade_2.0_ss_50sim_fixpoints_ensemblewise_synergies.tab")
ss_bliss_modelwise_50sim_file = paste0("results/link-only/bliss/cascade_2.0_ss_50sim_fixpoints_modelwise_synergies.tab")
ss_bliss_ensemblewise_100sim_file = paste0("results/link-only/bliss/cascade_2.0_ss_100sim_fixpoints_ensemblewise_synergies.tab")
ss_bliss_modelwise_100sim_file = paste0("results/link-only/bliss/cascade_2.0_ss_100sim_fixpoints_modelwise_synergies.tab")
ss_bliss_ensemblewise_150sim_file = paste0("results/link-only/bliss/cascade_2.0_ss_150sim_fixpoints_ensemblewise_synergies.tab")
ss_bliss_modelwise_150sim_file = paste0("results/link-only/bliss/cascade_2.0_ss_150sim_fixpoints_modelwise_synergies.tab")
ss_bliss_ensemblewise_200sim_file = paste0("results/link-only/bliss/cascade_2.0_ss_200sim_fixpoints_ensemblewise_synergies.tab")
ss_bliss_modelwise_200sim_file = paste0("results/link-only/bliss/cascade_2.0_ss_200sim_fixpoints_modelwise_synergies.tab")
prolif_bliss_ensemblewise_150sim_file = paste0("results/link-only/bliss/cascade_2.0_rand_150sim_fixpoints_bliss_ensemblewise_synergies.tab")
prolif_bliss_modelwise_150sim_file = paste0("results/link-only/bliss/cascade_2.0_rand_150sim_fixpoints_bliss_modelwise_synergies.tab")
random_bliss_ensemblewise_file = paste0("results/link-only/bliss/cascade_2.0_random_bliss_ensemblewise_synergies.tab")
random_bliss_modelwise_file = paste0("results/link-only/bliss/cascade_2.0_random_bliss_modelwise_synergies.tab")

ss_bliss_ensemblewise_synergies_50sim = emba::get_synergy_scores(ss_bliss_ensemblewise_50sim_file)
ss_bliss_modelwise_synergies_50sim = emba::get_synergy_scores(ss_bliss_modelwise_50sim_file, file_type = "modelwise")
ss_bliss_ensemblewise_synergies_100sim = emba::get_synergy_scores(ss_bliss_ensemblewise_100sim_file)
ss_bliss_modelwise_synergies_100sim = emba::get_synergy_scores(ss_bliss_modelwise_100sim_file, file_type = "modelwise")
ss_bliss_ensemblewise_synergies_150sim = emba::get_synergy_scores(ss_bliss_ensemblewise_150sim_file)
ss_bliss_modelwise_synergies_150sim = emba::get_synergy_scores(ss_bliss_modelwise_150sim_file, file_type = "modelwise")
ss_bliss_ensemblewise_synergies_200sim = emba::get_synergy_scores(ss_bliss_ensemblewise_200sim_file)
ss_bliss_modelwise_synergies_200sim = emba::get_synergy_scores(ss_bliss_modelwise_200sim_file, file_type = "modelwise")
prolif_bliss_ensemblewise_synergies_150sim = emba::get_synergy_scores(prolif_bliss_ensemblewise_150sim_file)
prolif_bliss_modelwise_synergies_150sim = emba::get_synergy_scores(prolif_bliss_modelwise_150sim_file, file_type = "modelwise")
random_bliss_ensemblewise_synergies = emba::get_synergy_scores(random_bliss_ensemblewise_file)
random_bliss_modelwise_synergies = emba::get_synergy_scores(random_bliss_modelwise_file, file_type = "modelwise")

# calculate probability of synergy in the modelwise results
ss_bliss_modelwise_synergies_50sim = ss_bliss_modelwise_synergies_50sim %>% 
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
ss_bliss_modelwise_synergies_100sim = ss_bliss_modelwise_synergies_100sim %>% 
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
ss_bliss_modelwise_synergies_150sim = ss_bliss_modelwise_synergies_150sim %>% 
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
ss_bliss_modelwise_synergies_200sim = ss_bliss_modelwise_synergies_200sim %>% 
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
prolif_bliss_modelwise_synergies_150sim = prolif_bliss_modelwise_synergies_150sim %>%
  mutate(synergy_prob_prolif = synergies/(synergies + `non-synergies`))
random_bliss_modelwise_synergies = random_bliss_modelwise_synergies %>%
  mutate(synergy_prob_random = synergies/(synergies + `non-synergies`))

# tidy data
pred_ew_bliss = bind_cols(
  ss_bliss_ensemblewise_synergies_50sim %>% select(perturbation, score) %>% rename(ss_score_50sim = score),
  ss_bliss_ensemblewise_synergies_100sim %>% select(score) %>% rename(ss_score_100sim = score),
  ss_bliss_ensemblewise_synergies_150sim %>% select(score) %>% rename(ss_score_150sim = score),
  ss_bliss_ensemblewise_synergies_200sim %>% select(score) %>% rename(ss_score_200sim = score),
  prolif_bliss_ensemblewise_synergies_150sim %>% select(score) %>% rename(prolif_score_150sim = score),
  random_bliss_ensemblewise_synergies %>% select(score) %>% rename(random_score = score), 
  as_tibble_col(observed, column_name = "observed"))

pred_mw_bliss = bind_cols(
  ss_bliss_modelwise_synergies_50sim %>% select(perturbation, synergy_prob_ss) %>% rename(synergy_prob_ss_50sim = synergy_prob_ss),
  ss_bliss_modelwise_synergies_100sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_ss_100sim = synergy_prob_ss),
  ss_bliss_modelwise_synergies_150sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_ss_150sim = synergy_prob_ss),
  ss_bliss_modelwise_synergies_200sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_ss_200sim = synergy_prob_ss),
  prolif_bliss_modelwise_synergies_150sim %>% select(synergy_prob_prolif) %>% rename(synergy_prob_prolif_150sim = synergy_prob_prolif),
  random_bliss_modelwise_synergies %>% select(synergy_prob_random),
  as_tibble_col(observed, column_name = "observed"))
```

### ROC curves {-}

```{r roc-bliss-cascade2, fig.width=5, fig.height=5, dpi=300, cache=TRUE, fig.show='hold', out.width='50%', fig.cap='ROC curves (CASCADE 2.0, Bliss synergy method)', fig.align='center'}
# 'ew' => ensemble-wise, 'mw' => model-wise
res_ss_ew_50sim = get_roc_stats(df = pred_ew_bliss, pred_col = "ss_score_50sim", label_col = "observed")
res_ss_ew_100sim = get_roc_stats(df = pred_ew_bliss, pred_col = "ss_score_100sim", label_col = "observed")
res_ss_ew_150sim = get_roc_stats(df = pred_ew_bliss, pred_col = "ss_score_150sim", label_col = "observed")
res_ss_ew_200sim = get_roc_stats(df = pred_ew_bliss, pred_col = "ss_score_200sim", label_col = "observed")
res_prolif_ew_150sim = get_roc_stats(df = pred_ew_bliss, pred_col = "prolif_score_150sim", label_col = "observed")
res_random_ew = get_roc_stats(df = pred_ew_bliss, pred_col = "random_score", label_col = "observed")

res_ss_mw_50sim = get_roc_stats(df = pred_mw_bliss, pred_col = "synergy_prob_ss_50sim", label_col = "observed", direction = ">")
res_ss_mw_100sim = get_roc_stats(df = pred_mw_bliss, pred_col = "synergy_prob_ss_100sim", label_col = "observed", direction = ">")
res_ss_mw_150sim = get_roc_stats(df = pred_mw_bliss, pred_col = "synergy_prob_ss_150sim", label_col = "observed", direction = ">")
res_ss_mw_200sim = get_roc_stats(df = pred_mw_bliss, pred_col = "synergy_prob_ss_200sim", label_col = "observed", direction = ">")
res_prolif_mw_150sim = get_roc_stats(df = pred_mw_bliss, pred_col = "synergy_prob_prolif_150sim", label_col = "observed", direction = ">")
res_random_mw = get_roc_stats(df = pred_mw_bliss, pred_col = "synergy_prob_random", label_col = "observed", direction = ">")

# Plot ROCs
plot(x = res_ss_ew_50sim$roc_stats$FPR, y = res_ss_ew_50sim$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Ensemble-wise synergies (Bliss)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = res_ss_ew_100sim$roc_stats$FPR, y = res_ss_ew_100sim$roc_stats$TPR,
  lwd = 3, col = my_palette[2])
lines(x = res_ss_ew_150sim$roc_stats$FPR, y = res_ss_ew_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[3])
lines(x = res_ss_ew_200sim$roc_stats$FPR, y = res_ss_ew_200sim$roc_stats$TPR,
  lwd = 3, col = my_palette[4])
lines(x = res_prolif_ew_150sim$roc_stats$FPR, y = res_prolif_ew_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[5])
lines(x = res_random_ew$roc_stats$FPR, y = res_random_ew$roc_stats$TPR, 
  lwd = 3, col = my_palette[6])
legend('topleft', title = 'AUC', col = my_palette[1:6], pch = 19,
  legend = c(paste(round(res_ss_ew_50sim$AUC, digits = 2), "Calibrated (50 sim)"),
    paste(round(res_ss_ew_100sim$AUC, digits = 2), "Calibrated (100 sim)"),
    paste(round(res_ss_ew_150sim$AUC, digits = 2), "Calibrated (150 sim)"),
    paste(round(res_ss_ew_200sim$AUC, digits = 2), "Calibrated (200 sim)"),
    paste(round(res_prolif_ew_150sim$AUC, digits = 2), "Proliferative (150 sim)"),
    paste(round(res_random_ew$AUC, digits = 2), "Random")), cex = 0.8)
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)

plot(x = res_ss_mw_50sim$roc_stats$FPR, y = res_ss_mw_50sim$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Model-wise synergies (Bliss)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = res_ss_mw_100sim$roc_stats$FPR, y = res_ss_mw_100sim$roc_stats$TPR,
  lwd = 3, col = my_palette[2])
lines(x = res_ss_mw_150sim$roc_stats$FPR, y = res_ss_mw_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[3])
lines(x = res_ss_mw_200sim$roc_stats$FPR, y = res_ss_mw_200sim$roc_stats$TPR,
  lwd = 3, col = my_palette[4])
lines(x = res_prolif_mw_150sim$roc_stats$FPR, y = res_prolif_mw_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[5])
lines(x = res_random_mw$roc_stats$FPR, y = res_random_mw$roc_stats$TPR, 
  lwd = 3, col = my_palette[6])
legend('bottomright', title = 'AUC', col = my_palette[1:6], pch = 19, cex = 0.9,
  legend = c(paste(round(res_ss_mw_50sim$AUC, digits = 2), "Calibrated (50 sim)"),
    paste(round(res_ss_mw_100sim$AUC, digits = 2), "Calibrated (100 sim)"),
    paste(round(res_ss_mw_150sim$AUC, digits = 2), "Calibrated (150 sim)"),
    paste(round(res_ss_mw_200sim$AUC, digits = 2), "Calibrated (200 sim)"),
    paste(round(res_prolif_mw_150sim$AUC, digits = 2), "Proliferative (150 sim)"),
    paste(round(res_random_mw$AUC, digits = 2), "Random")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)
```

### PR curves {-}

```{r pr-bliss-cascade2, fig.width=5, fig.height=5, dpi=300, cache=TRUE, fig.show='hold', out.width='50%', fig.cap='PR curves (CASCADE 2.0, Bliss synergy method)', fig.align='center'}
pr_ss_ew_bliss_50sim = pr.curve(scores.class0 = pred_ew_bliss %>% pull(ss_score_50sim) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_bliss %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_ss_ew_bliss_100sim = pr.curve(scores.class0 = pred_ew_bliss %>% pull(ss_score_100sim) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_bliss %>% pull(observed), curve = TRUE)
pr_ss_ew_bliss_150sim = pr.curve(scores.class0 = pred_ew_bliss %>% pull(ss_score_150sim) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_bliss %>% pull(observed), curve = TRUE)
pr_ss_ew_bliss_200sim = pr.curve(scores.class0 = pred_ew_bliss %>% pull(ss_score_200sim) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_bliss %>% pull(observed), curve = TRUE)
pr_prolif_ew_bliss_150sim = pr.curve(scores.class0 = pred_ew_bliss %>% pull(prolif_score_150sim) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_bliss %>% pull(observed), curve = TRUE)
pr_random_ew_bliss = pr.curve(scores.class0 = pred_ew_bliss %>% pull(random_score) %>% (function(x) {-x}), 
  weights.class0 = pred_ew_bliss %>% pull(observed), curve = TRUE)

pr_ss_mw_bliss_50sim = pr.curve(scores.class0 = pred_mw_bliss %>% pull(synergy_prob_ss_50sim), 
  weights.class0 = pred_mw_bliss %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_ss_mw_bliss_100sim = pr.curve(scores.class0 = pred_mw_bliss %>% pull(synergy_prob_ss_100sim), 
  weights.class0 = pred_mw_bliss %>% pull(observed), curve = TRUE)
pr_ss_mw_bliss_150sim = pr.curve(scores.class0 = pred_mw_bliss %>% pull(synergy_prob_ss_150sim), 
  weights.class0 = pred_mw_bliss %>% pull(observed), curve = TRUE)
pr_ss_mw_bliss_200sim = pr.curve(scores.class0 = pred_mw_bliss %>% pull(synergy_prob_ss_200sim), 
  weights.class0 = pred_mw_bliss %>% pull(observed), curve = TRUE)
pr_prolif_mw_bliss_150sim = pr.curve(scores.class0 = pred_mw_bliss %>% pull(synergy_prob_prolif_150sim), 
  weights.class0 = pred_mw_bliss %>% pull(observed), curve = TRUE)
pr_random_mw_bliss = pr.curve(scores.class0 = pred_mw_bliss %>% pull(synergy_prob_random), 
  weights.class0 = pred_mw_bliss %>% pull(observed), curve = TRUE)

plot(pr_ss_ew_bliss_50sim, main = 'PR curve, Ensemble-wise synergies (Bliss)',
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_ss_ew_bliss_100sim, add = TRUE, color = my_palette[2])
plot(pr_ss_ew_bliss_150sim, add = TRUE, color = my_palette[3])
plot(pr_ss_ew_bliss_200sim, add = TRUE, color = my_palette[4])
plot(pr_prolif_ew_bliss_150sim, add = TRUE, color = my_palette[5])
plot(pr_random_ew_bliss, add = TRUE, color = my_palette[6])
legend('topright', title = 'AUC', col = my_palette[1:6], pch = 19,
  legend = c(paste(round(pr_ss_ew_bliss_50sim$auc.davis.goadrich, digits = 3), "Calibrated (50 sim)"),
    paste(round(pr_ss_ew_bliss_100sim$auc.davis.goadrich, digits = 3), "Calibrated (100 sim)"),
    paste(round(pr_ss_ew_bliss_150sim$auc.davis.goadrich, digits = 3), "Calibrated (150 sim)"),
    paste(round(pr_ss_ew_bliss_200sim$auc.davis.goadrich, digits = 3), "Calibrated (200 sim)"),
    paste(round(pr_prolif_ew_bliss_150sim$auc.davis.goadrich, digits = 3), "Proliferative (150 sim)"),
    paste(round(pr_random_ew_bliss$auc.davis.goadrich, digits = 3), "Random")))
grid(lwd = 0.5)

plot(pr_ss_mw_bliss_50sim, main = 'PR curve, Model-wise synergies (Bliss)', 
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_ss_mw_bliss_100sim, add = TRUE, color = my_palette[2])
plot(pr_ss_mw_bliss_150sim, add = TRUE, color = my_palette[3])
plot(pr_ss_mw_bliss_200sim, add = TRUE, color = my_palette[4])
plot(pr_prolif_mw_bliss_150sim, add = TRUE, color = my_palette[5])
plot(pr_random_mw_bliss, add = TRUE, color = my_palette[6])
legend('topright', title = 'AUC', col = my_palette[1:6], pch = 19,
  legend = c(paste(round(pr_ss_mw_bliss_50sim$auc.davis.goadrich, digits = 3), "Calibrated (50 sim)"),
    paste(round(pr_ss_mw_bliss_100sim$auc.davis.goadrich, digits = 3), "Calibrated (100 sim)"),
    paste(round(pr_ss_mw_bliss_150sim$auc.davis.goadrich, digits = 3), "Calibrated (150 sim)"),
    paste(round(pr_ss_mw_bliss_200sim$auc.davis.goadrich, digits = 3), "Calibrated (200 sim)"),
    paste(round(pr_prolif_mw_bliss_150sim$auc.davis.goadrich, digits = 3), "Proliferative (150 sim)"),
    paste(round(pr_random_mw_bliss$auc.davis.goadrich, digits = 3), "Random")))
grid(lwd = 0.5)
```

:::{.green-box}
- To minimize the resulting performance variance, $150$ seems to be a good number of `Gitsbe` simulations to run for the CASCADE 2.0 network.
- Individual predictor **model-wise** results (when looking at the ROC curves) show good performance.
- Individual predictor **ensemble-wise** results show that *proliferative* and *calibrated* models have poor performance whereas *random* models perform like proper random models ($AUC\sim0.5$))
- The PR curves show that the **performance of all individual predictors is poor** compared to the baseline.
:::

### AUC sensitivity {-}

Investigate same thing as described in [here](#auc-sensitivity).
This is very crucial since the PR performance is poor for the individual predictors and the ensemble-wise predictors were really bad in terms for AUC-ROC, but a combined predictor might be able to counter this.
We will combine the synergy scores from the *random* and *proliferative* simulations with the results from the *calibrated* Gitsbe simulations ($150$).

```{r auc-sen-ew-bliss-cascade2, dpi=300, fig.align='center', cache=TRUE, fig.cap='AUC sensitivity (CASCADE 2.0, Bliss synergy method, Ensemble-wise results)'}
# Ensemble-wise
betas = seq(from = -10, to = 10, by = 0.1)

random_roc = sapply(betas, function(beta) {
  pred_ew_bliss = pred_ew_bliss %>% mutate(combined_score = ss_score_150sim + beta * random_score)
  res = roc.curve(scores.class0 = pred_ew_bliss %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_bliss %>% pull(observed))
  auc_value = res$auc
})

prolif_roc = sapply(betas, function(beta) {
  pred_ew_bliss = pred_ew_bliss %>% mutate(combined_score = ss_score_150sim + beta * prolif_score_150sim)
  res = roc.curve(scores.class0 = pred_ew_bliss %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_bliss %>% pull(observed))
  auc_value = res$auc
})

random_pr = sapply(betas, function(beta) {
  pred_ew_bliss = pred_ew_bliss %>% mutate(combined_score = ss_score_150sim + beta * random_score)
  res = pr.curve(scores.class0 = pred_ew_bliss %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_bliss %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

prolif_pr = sapply(betas, function(beta) {
  pred_ew_bliss = pred_ew_bliss %>% mutate(combined_score = ss_score_150sim + beta * prolif_score_150sim)
  res = pr.curve(scores.class0 = pred_ew_bliss %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_ew_bliss %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_ew = as_tibble(cbind(betas, random_roc, prolif_roc, random_pr, prolif_pr))
df_ew = df_ew %>% tidyr::pivot_longer(-betas, names_to = "type", values_to = "AUC")

ggline(data = df_ew, x = "betas", y = "AUC", numeric.x.axis = TRUE, color = "type",
  plot_type = "l", xlab = TeX("$\\beta$"), ylab = "AUC (Area Under Curve)", 
  legend = "none", facet.by = "type", palette = my_palette,
  panel.labs = list(type = c("PR: calibrated + β x proliferative", 
    "ROC: calibrated + β x proliferative", "PR: calibrated + β x random", 
    "ROC: calibrated + β x random")),
  title = TeX("AUC sensitivity to $\\beta$ parameter (Bliss, CASCADE 2.0)")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = -1.6, color = "red", size = 0.3, linetype = "dashed") + 
  geom_text(aes(x=-4, label="β = -1.6", y=0.15), colour="black") + 
  grids()
```

```{r auc-sen-mw-bliss-cascade2, dpi=300, fig.align='center', cache=TRUE, fig.cap='AUC sensitivity (CASCADE 2.0, Bliss synergy method, Model-wise results)'}
# Model-wise
weights = seq(from = 0, to = 1, by = 0.05)

random_roc_mw = sapply(weights, function(w) {
  pred_mw_bliss = pred_mw_bliss %>%
    mutate(weighted_prob = (1 - w) * pred_mw_bliss$synergy_prob_ss_150sim + w * pred_mw_bliss$synergy_prob_random)
  res = roc.curve(scores.class0 = pred_mw_bliss %>% pull(weighted_prob),
    weights.class0 = pred_mw_bliss %>% pull(observed))
  auc_value = res$auc
})

prolif_roc_mw = sapply(weights, function(w) {
  pred_mw_bliss = pred_mw_bliss %>%
    mutate(weighted_prob = (1 - w) * pred_mw_bliss$synergy_prob_ss_150sim + w * pred_mw_bliss$synergy_prob_prolif_150sim)
  res = roc.curve(scores.class0 = pred_mw_bliss %>% pull(weighted_prob),
    weights.class0 = pred_mw_bliss %>% pull(observed))
  auc_value = res$auc
})

random_pr_mw = sapply(weights, function(w) {
  pred_mw_bliss = pred_mw_bliss %>% 
    mutate(weighted_prob = (1 - w) * pred_mw_bliss$synergy_prob_ss_150sim + w * pred_mw_bliss$synergy_prob_random)
  res = pr.curve(scores.class0 = pred_mw_bliss %>% pull(weighted_prob), 
    weights.class0 = pred_mw_bliss %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

prolif_pr_mw = sapply(weights, function(w) {
  pred_mw_bliss = pred_mw_bliss %>% 
    mutate(weighted_prob = (1 - w) * pred_mw_bliss$synergy_prob_ss_150sim + w * pred_mw_bliss$synergy_prob_prolif_150sim)
  res = pr.curve(scores.class0 = pred_mw_bliss %>% pull(weighted_prob), 
    weights.class0 = pred_mw_bliss %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_mw = as_tibble(cbind(weights, random_roc_mw, prolif_roc_mw, random_pr_mw, prolif_pr_mw))
df_mw = df_mw %>% tidyr::pivot_longer(-weights, names_to = "type", values_to = "AUC")

ggline(data = df_mw, x = "weights", y = "AUC", numeric.x.axis = TRUE, color = "type",
  plot_type = "l", xlab = TeX("weight $w$"), ylab = "AUC (Area Under Curve)", 
  legend = "none", facet.by = "type", palette = my_palette,
  panel.labs = list(type = c("PR: (1-w) x prob(ss) + w x prob(prolif)", 
    "ROC: (1-w) x prob(ss) + w x prob(prolif)", "PR: (1-w) x prob(ss) + w x prob(random)", 
    "ROC: (1-w) x prob(ss) + w x prob(random)")), title.position = "center",
  title = TeX("AUC sensitivity to weighted average score (Bliss, CASCADE 2.0)")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  grids()
```

:::{.green-box}
- No added benefit when using the *model-wise* approach.
- The random models do not augment the prediction performance of the calibrated models at all.
- The proliferative models can be used to normalize against the predictions of the calibrated models and thus bring significant contribution to the calibrated models performance (both ROC-AUC and PR-AUC are increased).
- The $\beta_{best}$ that maximizes the ROC and PR AUC for the **combination of proliferative and calibrated models** and is equal to $\beta_{best}=-1.6$.
For $\beta=-1$ we still see **significant performance improvement**.
:::

## Correlation {-}

We test for correlation between some of the results shown in the ROC curves.
The results tested are the *ensemble-wise* vs *model-wise*, *random* models vs *calibrated (ss)* models and *HSA* vs *Bliss* synergy assessment (the calibrated and proliferative models are from the $150$ simulation results).
*P-values* are represented at 3 significant levels: $0.05, 0.01, 0.001$ (\*, \*\*, \*\*\*) and the correlation coefficient is calculated using Kendall's *tau* statistic.

```{r cor-plot-cascade2, dpi=300, warning=FALSE, cache=TRUE, fig.align='center', fig.cap='Correlation Plot for CASCADE 2.0 Results'}
synergy_scores = bind_cols(
  pred_ew_hsa %>% select(ss_score_150sim, prolif_score_150sim, random_score) %>% rename(ss_ew_hsa = ss_score_150sim, prolif_ew_hsa = prolif_score_150sim, random_ew_hsa = random_score),
  pred_ew_bliss %>% select(ss_score_150sim, prolif_score_150sim, random_score) %>% rename(ss_ew_bliss = ss_score_150sim, prolif_ew_bliss = prolif_score_150sim, random_ew_bliss = random_score),
  pred_mw_hsa %>% select(synergy_prob_ss_150sim, synergy_prob_prolif_150sim, synergy_prob_random) %>% 
    rename(ss_mw_hsa = synergy_prob_ss_150sim, prolif_mw_hsa = synergy_prob_prolif_150sim, random_mw_hsa = synergy_prob_random),
  pred_mw_bliss %>% select(synergy_prob_ss_150sim, synergy_prob_prolif_150sim, synergy_prob_random) %>% 
    rename(ss_mw_bliss = synergy_prob_ss_150sim, prolif_mw_bliss = synergy_prob_prolif_150sim, random_mw_bliss = synergy_prob_random)
  )

M = cor(synergy_scores, method = "kendall")
res = cor.mtest(synergy_scores, method = "kendall")
corrplot(corr = M, type = "upper", p.mat = res$p, sig.level = c(.001, .01, .05), 
  pch.cex = 1, pch.col = "white", insig = "label_sig", tl.col = "black", tl.srt = 45)
```

:::{.green-box}
- **Bliss ensemble-wise** results don't correlate at all with the **model-wise** results (*topright* part of the correlation plot).
The **HSA ensemble-wise** results do so (at some degree).
- Between the **ensemble-wise** results there is no strong correlation (*topleft*) while between the **model-wise** (*bottomright*) there is strong correlation.
- **Ensemble-wise calibrated** results seem to correlate more with the **proliferative** than with the **random results** (*topleft*).
:::

## Fitness vs Performance {-}

The idea here is to generate many training data files from the steady state as used in the simulations above, where some of the nodes will have their states *flipped* to the opposite state ($0$ to $1$ and vice versa).
That way, we can train models to different steady states, ranging from ones that differ to just a few nodes states up to a steady state that is the complete *reversed* version of the one used in the simulations.

Using the [gen_training_data.R](https://github.com/bblodfon/ags-paper-1/blob/master/scripts/gen_training_data.R) script, we first choose a few number of flips ($11$ flips) ranging from $1$ to $24$ (all nodes) in the steady state.
Then, for each such *flipping-nodes* value, we generated $20$ new steady states with a randomly chosen set of nodes whose value is going to flip.
Thus, in total, $205$ training data files were produced ($205 = 9 \times 20 + 24 + 1$, where from the $11$ number of flips, the one flip happens for every node and flipping all the nodes simultaneously happens once).

Running the script [run_druglogics_synergy_training.sh](https://github.com/bblodfon/ags-paper-1/blob/master/scripts/run_druglogics_synergy_training.sh) from the `druglogics-synergy` repository root, we get the simulation results for each of these training data files.
Note that in the CASCADE 2.0 configuration file we changed the number of simulations to ($15$) for each training data file, the attractor tool used was `biolqm_stable_states` (as is with every analysis in this report) and the `synergy_method: hsa`.

We now load the data from these simulations:

# CASCADE 2.0 Analysis (Topology Mutations) {-}

:::{.note}
We run `Gitsbe` simulations with $50$ topology mutations (bootstrap value, reduced to $10$ after models with stabla states have been found), both for $50$ and $150$ simulations and both **fitting to steady state** (calibrated models) and to a **proliferative phenotype** (so not random models but as close as it can get to that since we are discussing topology mutations).

Results were assessed via `Drabme`'s ensemble and modelwise results, using both the HSA and Bliss method.
:::

Load the results:

```{r Load Topology Mutation results}
# 'ss' => calibrated models, 'rand' => proliferative models (so not random but kind of!)
# 'ew' => ensemble-wise, 'mw' => modelwise

## HSA results ss
topo_ss_hsa_ew_50sim_file = paste0("results/topology-only/cascade_2.0_ss_50sim_fixpoints_hsa_ensemblewise_synergies.tab")
topo_ss_hsa_mw_50sim_file = paste0("results/topology-only/cascade_2.0_ss_50sim_fixpoints_hsa_modelwise_synergies.tab")
topo_ss_hsa_ew_150sim_file = paste0("results/topology-only/cascade_2.0_ss_150sim_fixpoints_hsa_ensemblewise_synergies.tab")
topo_ss_hsa_mw_150sim_file = paste0("results/topology-only/cascade_2.0_ss_150sim_fixpoints_hsa_modelwise_synergies.tab")

topo_ss_hsa_ew_synergies_50sim = emba::get_synergy_scores(topo_ss_hsa_ew_50sim_file)
topo_ss_hsa_mw_synergies_50sim = emba::get_synergy_scores(topo_ss_hsa_mw_50sim_file, file_type = "modelwise")
topo_ss_hsa_ew_synergies_150sim = emba::get_synergy_scores(topo_ss_hsa_ew_150sim_file)
topo_ss_hsa_mw_synergies_150sim = emba::get_synergy_scores(topo_ss_hsa_mw_150sim_file, file_type = "modelwise")

## HSA results rand
topo_prolif_hsa_ew_50sim_file = paste0("results/topology-only/cascade_2.0_rand_50sim_fixpoints_hsa_ensemblewise_synergies.tab")
topo_prolif_hsa_mw_50sim_file = paste0("results/topology-only/cascade_2.0_rand_50sim_fixpoints_hsa_modelwise_synergies.tab")
topo_prolif_hsa_ew_150sim_file = paste0("results/topology-only/cascade_2.0_rand_150sim_fixpoints_hsa_ensemblewise_synergies.tab")
topo_prolif_hsa_mw_150sim_file = paste0("results/topology-only/cascade_2.0_rand_150sim_fixpoints_hsa_modelwise_synergies.tab")

topo_prolif_hsa_ew_synergies_50sim = emba::get_synergy_scores(topo_prolif_hsa_ew_50sim_file)
topo_prolif_hsa_mw_synergies_50sim = emba::get_synergy_scores(topo_prolif_hsa_mw_50sim_file, file_type = "modelwise")
topo_prolif_hsa_ew_synergies_150sim = emba::get_synergy_scores(topo_prolif_hsa_ew_150sim_file)
topo_prolif_hsa_mw_synergies_150sim = emba::get_synergy_scores(topo_prolif_hsa_mw_150sim_file, file_type = "modelwise")

## Bliss results ss
topo_ss_bliss_ew_50sim_file = paste0("results/topology-only/cascade_2.0_ss_50sim_fixpoints_bliss_ensemblewise_synergies.tab")
topo_ss_bliss_mw_50sim_file = paste0("results/topology-only/cascade_2.0_ss_50sim_fixpoints_bliss_modelwise_synergies.tab")
topo_ss_bliss_ew_150sim_file = paste0("results/topology-only/cascade_2.0_ss_150sim_fixpoints_bliss_ensemblewise_synergies.tab")
topo_ss_bliss_mw_150sim_file = paste0("results/topology-only/cascade_2.0_ss_150sim_fixpoints_bliss_modelwise_synergies.tab")

topo_ss_bliss_ew_synergies_50sim = emba::get_synergy_scores(topo_ss_bliss_ew_50sim_file)
topo_ss_bliss_mw_synergies_50sim = emba::get_synergy_scores(topo_ss_bliss_mw_50sim_file, file_type = "modelwise")
topo_ss_bliss_ew_synergies_150sim = emba::get_synergy_scores(topo_ss_bliss_ew_150sim_file)
topo_ss_bliss_mw_synergies_150sim = emba::get_synergy_scores(topo_ss_bliss_mw_150sim_file, file_type = "modelwise")

## Bliss results rand
topo_prolif_bliss_ew_50sim_file = paste0("results/topology-only/cascade_2.0_rand_50sim_fixpoints_bliss_ensemblewise_synergies.tab")
topo_prolif_bliss_mw_50sim_file = paste0("results/topology-only/cascade_2.0_rand_50sim_fixpoints_bliss_modelwise_synergies.tab")
topo_prolif_bliss_ew_150sim_file = paste0("results/topology-only/cascade_2.0_rand_150sim_fixpoints_bliss_ensemblewise_synergies.tab")
topo_prolif_bliss_mw_150sim_file = paste0("results/topology-only/cascade_2.0_rand_150sim_fixpoints_bliss_modelwise_synergies.tab")

topo_prolif_bliss_ew_synergies_50sim = emba::get_synergy_scores(topo_prolif_bliss_ew_50sim_file)
topo_prolif_bliss_mw_synergies_50sim = emba::get_synergy_scores(topo_prolif_bliss_mw_50sim_file, file_type = "modelwise")
topo_prolif_bliss_ew_synergies_150sim = emba::get_synergy_scores(topo_prolif_bliss_ew_150sim_file)
topo_prolif_bliss_mw_synergies_150sim = emba::get_synergy_scores(topo_prolif_bliss_mw_150sim_file, file_type = "modelwise")

# calculate probability of synergy in the modelwise results
topo_ss_hsa_mw_synergies_50sim = topo_ss_hsa_mw_synergies_50sim %>% 
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topo_ss_hsa_mw_synergies_150sim = topo_ss_hsa_mw_synergies_150sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topo_prolif_hsa_mw_synergies_50sim = topo_prolif_hsa_mw_synergies_50sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topo_prolif_hsa_mw_synergies_150sim = topo_prolif_hsa_mw_synergies_150sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topo_ss_bliss_mw_synergies_50sim = topo_ss_bliss_mw_synergies_50sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topo_ss_bliss_mw_synergies_150sim = topo_ss_bliss_mw_synergies_150sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topo_prolif_bliss_mw_synergies_50sim = topo_prolif_bliss_mw_synergies_50sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topo_prolif_bliss_mw_synergies_150sim = topo_prolif_bliss_mw_synergies_150sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))

# Tidy the data
pred_topo_ew_hsa = bind_cols(
  topo_ss_hsa_ew_synergies_50sim %>% rename(ss_score_50sim = score),
  topo_ss_hsa_ew_synergies_150sim %>% select(score) %>% rename(ss_score_150sim = score),
  topo_prolif_hsa_ew_synergies_50sim %>% select(score) %>% rename(prolif_score_50sim = score),
  topo_prolif_hsa_ew_synergies_150sim %>% select(score) %>% rename(prolif_score_150sim = score),
  as_tibble_col(observed, column_name = "observed"))

pred_topo_mw_hsa = bind_cols(
  topo_ss_hsa_mw_synergies_50sim %>% select(perturbation, synergy_prob_ss) %>% rename(synergy_prob_ss_50sim = synergy_prob_ss),
  topo_ss_hsa_mw_synergies_150sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_ss_150sim = synergy_prob_ss),
  topo_prolif_hsa_mw_synergies_50sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_prolif_50sim = synergy_prob_ss),
  topo_prolif_hsa_mw_synergies_150sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_prolif_150sim = synergy_prob_ss),
  as_tibble_col(observed, column_name = "observed"))

pred_topo_ew_bliss = bind_cols(
  topo_ss_bliss_ew_synergies_50sim %>% rename(ss_score_50sim = score),
  topo_ss_bliss_ew_synergies_150sim %>% select(score) %>% rename(ss_score_150sim = score),
  topo_prolif_bliss_ew_synergies_50sim %>% select(score) %>% rename(prolif_score_50sim = score),
  topo_prolif_bliss_ew_synergies_150sim %>% select(score) %>% rename(prolif_score_150sim = score),
  as_tibble_col(observed, column_name = "observed"))

pred_topo_mw_bliss = bind_cols(
  topo_ss_bliss_mw_synergies_50sim %>% select(perturbation, synergy_prob_ss) %>% rename(synergy_prob_ss_50sim = synergy_prob_ss),
  topo_ss_bliss_mw_synergies_150sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_ss_150sim = synergy_prob_ss),
  topo_prolif_bliss_mw_synergies_50sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_prolif_50sim = synergy_prob_ss),
  topo_prolif_bliss_mw_synergies_150sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_prolif_150sim = synergy_prob_ss),
  as_tibble_col(observed, column_name = "observed"))
```

## ROC curves (HSA) {-}

```{r ROC curves Topology Mutations (CASCADE 2.0 - HSA), fig.width=6, fig.height=6, dpi=300, cache=TRUE, fig.show='hold', out.width='50%'}
topo_res_ss_ew_50sim = get_roc_stats(df = pred_topo_ew_hsa, pred_col = "ss_score_50sim", label_col = "observed")
topo_res_ss_ew_150sim = get_roc_stats(df = pred_topo_ew_hsa, pred_col = "ss_score_150sim", label_col = "observed")
topo_res_prolif_ew_50sim = get_roc_stats(df = pred_topo_ew_hsa, pred_col = "prolif_score_50sim", label_col = "observed")
topo_res_prolif_ew_150sim = get_roc_stats(df = pred_topo_ew_hsa, pred_col = "prolif_score_150sim", label_col = "observed")

topo_res_ss_mw_50sim = get_roc_stats(df = pred_topo_mw_hsa, pred_col = "synergy_prob_ss_50sim", label_col = "observed", direction = ">")
topo_res_ss_mw_150sim = get_roc_stats(df = pred_topo_mw_hsa, pred_col = "synergy_prob_ss_150sim", label_col = "observed", direction = ">")
topo_res_prolif_mw_50sim = get_roc_stats(df = pred_topo_mw_hsa, pred_col = "synergy_prob_prolif_50sim", label_col = "observed", direction = ">")
topo_res_prolif_mw_150sim = get_roc_stats(df = pred_topo_mw_hsa, pred_col = "synergy_prob_prolif_150sim", label_col = "observed", direction = ">")

# Plot ROCs
plot(x = topo_res_ss_ew_50sim$roc_stats$FPR, y = topo_res_ss_ew_50sim$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Ensemble-wise synergies (HSA)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = topo_res_ss_ew_150sim$roc_stats$FPR, y = topo_res_ss_ew_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[2])
lines(x = topo_res_prolif_ew_50sim$roc_stats$FPR, y = topo_res_prolif_ew_50sim$roc_stats$TPR,
  lwd = 3, col = my_palette[3])
lines(x = topo_res_prolif_ew_150sim$roc_stats$FPR, y = topo_res_prolif_ew_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[4])
legend('bottomright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(topo_res_ss_ew_50sim$AUC, digits = 2), "Calibrated (50 sim)"),
    paste(round(topo_res_ss_ew_150sim$AUC, digits = 2), "Calibrated (150 sim)"),
    paste(round(topo_res_prolif_ew_50sim$AUC, digits = 2), "Proliferative (50 sim)"),
    paste(round(topo_res_prolif_ew_150sim$AUC, digits = 2), "Proliferative (150 sim)")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)

plot(x = topo_res_ss_mw_50sim$roc_stats$FPR, y = topo_res_ss_mw_50sim$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Model-wise synergies (HSA)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = topo_res_ss_mw_150sim$roc_stats$FPR, y = topo_res_ss_mw_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[2])
lines(x = topo_res_prolif_mw_50sim$roc_stats$FPR, y = topo_res_prolif_mw_50sim$roc_stats$TPR,
  lwd = 3, col = my_palette[3])
lines(x = topo_res_prolif_mw_150sim$roc_stats$FPR, y = topo_res_prolif_mw_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[4])
legend('bottomright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(topo_res_ss_mw_50sim$AUC, digits = 2), "Calibrated (50 sim)"),
    paste(round(topo_res_ss_mw_150sim$AUC, digits = 2), "Calibrated (150 sim)"),
    paste(round(topo_res_prolif_mw_50sim$AUC, digits = 2), "Proliferative (50 sim)"),
    paste(round(topo_res_prolif_mw_150sim$AUC, digits = 2), "Proliferative (150 sim)")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)
```

## PR curves (HSA) {-}

```{r PR curves Topology Mutations (CASCADE 2.0 - HSA), fig.width=6, fig.height=6, dpi=300, cache=TRUE, fig.show='hold', out.width='50%'}
pr_topo_res_ss_ew_50sim = pr.curve(scores.class0 = pred_topo_ew_hsa %>% pull(ss_score_50sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topo_ew_hsa %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_topo_res_ss_ew_150sim = pr.curve(scores.class0 = pred_topo_ew_hsa %>% pull(ss_score_150sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topo_ew_hsa %>% pull(observed), curve = TRUE)
pr_topo_res_prolif_ew_50sim = pr.curve(scores.class0 = pred_topo_ew_hsa %>% pull(prolif_score_50sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topo_ew_hsa %>% pull(observed), curve = TRUE)
pr_topo_res_prolif_ew_150sim = pr.curve(scores.class0 = pred_topo_ew_hsa %>% pull(prolif_score_150sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topo_ew_hsa %>% pull(observed), curve = TRUE)

pr_topo_res_ss_mw_50sim = pr.curve(scores.class0 = pred_topo_mw_hsa %>% pull(synergy_prob_ss_50sim),
  weights.class0 = pred_topo_mw_hsa %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_topo_res_ss_mw_150sim = pr.curve(scores.class0 = pred_topo_mw_hsa %>% pull(synergy_prob_ss_150sim),
  weights.class0 = pred_topo_mw_hsa %>% pull(observed), curve = TRUE)
pr_topo_res_prolif_mw_50sim = pr.curve(scores.class0 = pred_topo_mw_hsa %>% pull(synergy_prob_prolif_50sim),
  weights.class0 = pred_topo_mw_hsa %>% pull(observed), curve = TRUE)
pr_topo_res_prolif_mw_150sim = pr.curve(scores.class0 = pred_topo_mw_hsa %>% pull(synergy_prob_prolif_150sim),
  weights.class0 = pred_topo_mw_hsa %>% pull(observed), curve = TRUE)

plot(pr_topo_res_ss_ew_50sim, main = 'PR curve, Ensemble-wise synergies (HSA)',
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_topo_res_ss_ew_150sim, add = TRUE, color = my_palette[2])
plot(pr_topo_res_prolif_ew_50sim, add = TRUE, color = my_palette[3])
plot(pr_topo_res_prolif_ew_150sim, add = TRUE, color = my_palette[4])
legend('topright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(pr_topo_res_ss_ew_50sim$auc.davis.goadrich, digits = 3), "Calibrated (50 sim)"),
    paste(round(pr_topo_res_ss_ew_150sim$auc.davis.goadrich, digits = 3), "Calibrated (150 sim)"),
    paste(round(pr_topo_res_prolif_ew_50sim$auc.davis.goadrich, digits = 3), "Proliferative (50 sim)"),
    paste(round(pr_topo_res_prolif_ew_150sim$auc.davis.goadrich, digits = 3), "Proliferative (150 sim)")))
grid(lwd = 0.5)

plot(pr_topo_res_ss_mw_50sim, main = 'PR curve, Model-wise synergies (HSA)',
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_topo_res_ss_mw_150sim, add = TRUE, color = my_palette[2])
plot(pr_topo_res_prolif_mw_50sim, add = TRUE, color = my_palette[3])
plot(pr_topo_res_prolif_mw_150sim, add = TRUE, color = my_palette[4])
legend('topright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(pr_topo_res_ss_mw_50sim$auc.davis.goadrich, digits = 3), "Calibrated (50 sim)"),
    paste(round(pr_topo_res_ss_mw_150sim$auc.davis.goadrich, digits = 3), "Calibrated (150 sim)"),
    paste(round(pr_topo_res_prolif_mw_50sim$auc.davis.goadrich, digits = 3), "Proliferative (50 sim)"),
    paste(round(pr_topo_res_prolif_mw_150sim$auc.davis.goadrich, digits = 3), "Proliferative (150 sim)")))
grid(lwd = 0.5)
```

## ROC-AUC sensitivity (HSA) {-}

Combine the $150$ simulation results (calibrated + proliferative)
```{r ROC-AUC sensitivity Topology Mutations (HSA - CASCADE 2.0), dpi=300, out.width='80%', fig.align='center', cache=TRUE}
# Ensemble-wise
betas = seq(from = -10, to = 10, by = 0.1)

auc_values_ew = sapply(betas, function(beta) {
  pred_topo_ew_hsa = pred_topo_ew_hsa %>% mutate(combined_score = ss_score_150sim + beta * prolif_score_150sim)
  res = get_roc_stats(df = pred_topo_ew_hsa, pred_col = "combined_score", label_col = "observed")
  auc_value = res$AUC
})

df_ew = as_tibble(cbind(betas, auc_values_ew))

ggline(data = df_ew, x = "betas", y = "auc_values_ew", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("$\\beta$"), ylab = "AUC (Area Under ROC Curve)",
  title = TeX("AUC sensitivity to $\\beta$ parameter: $calibrated + \\beta \\times random$"),
  color = my_palette[2]) + geom_vline(xintercept = 0) + grids()

# Model-wise
weights = seq(from = 0, to = 1, by = 0.05)

auc_values_mw = sapply(weights, function(w) {
  pred_topo_mw_hsa = pred_topo_mw_hsa %>% 
    mutate(weighted_prob = (1 - w) * pred_topo_mw_hsa$synergy_prob_ss_150sim + w * pred_topo_mw_hsa$synergy_prob_prolif_150sim)
  res = get_roc_stats(df = pred_topo_mw_hsa, pred_col = "weighted_prob", label_col = "observed", direction = ">")
  auc_value = res$AUC
})

df_mw = as_tibble(cbind(weights, auc_values_mw))

ggline(data = df_mw, x = "weights", y = "auc_values_mw", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("weight $w$"), ylab = "AUC (Area Under ROC Curve)",
  title = TeX("AUC sensitivity to weighted average score: $(1-w) \\times prob_{ss} + w \\times prob_{rand}$"),
  color = my_palette[3]) + grids()
```

## PR-AUC sensitivity (HSA) {-}

Combine the $150$ simulation results (calibrated + proliferative)
```{r PR-AUC sensitivity Topology Mutations (HSA - CASCADE 2.0), dpi=300, out.width='80%', fig.align='center', cache=TRUE}
# Ensemble-wise
betas = seq(from = -10, to = 10, by = 0.1)

auc_values_ew = sapply(betas, function(beta) {
  pred_topo_ew_hsa = pred_topo_ew_hsa %>% mutate(combined_score = ss_score_150sim + beta * prolif_score_150sim)
  res = pr.curve(scores.class0 = pred_topo_ew_hsa %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_topo_ew_hsa %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_ew = as_tibble(cbind(betas, auc_values_ew))

ggline(data = df_ew, x = "betas", y = "auc_values_ew", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("$\\beta$"), ylab = "AUC (Area Under PR Curve)",
  title = TeX("AUC sensitivity to $\\beta$ parameter: $calibrated + \\beta \\times random$"),
  color = my_palette[2], label = "betas", repel = TRUE,
  label.select = list(criteria = "`y` == max(`y`)"),
  font.label = list(size = 14)) + geom_vline(xintercept = 0) + grids()
```

## ROC curves (Bliss) {-}

```{r ROC curves Topology Mutations (CASCADE 2.0 - Bliss), fig.width=6, fig.height=6, dpi=300, cache=TRUE, fig.show='hold', out.width='50%'}
topo_res_ss_ew_50sim = get_roc_stats(df = pred_topo_ew_bliss, pred_col = "ss_score_50sim", label_col = "observed")
topo_res_ss_ew_150sim = get_roc_stats(df = pred_topo_ew_bliss, pred_col = "ss_score_150sim", label_col = "observed")
topo_res_prolif_ew_50sim = get_roc_stats(df = pred_topo_ew_bliss, pred_col = "prolif_score_50sim", label_col = "observed")
topo_res_prolif_ew_150sim = get_roc_stats(df = pred_topo_ew_bliss, pred_col = "prolif_score_150sim", label_col = "observed")

topo_res_ss_mw_50sim = get_roc_stats(df = pred_topo_mw_bliss, pred_col = "synergy_prob_ss_50sim", label_col = "observed", direction = ">")
topo_res_ss_mw_150sim = get_roc_stats(df = pred_topo_mw_bliss, pred_col = "synergy_prob_ss_150sim", label_col = "observed", direction = ">")
topo_res_prolif_mw_50sim = get_roc_stats(df = pred_topo_mw_bliss, pred_col = "synergy_prob_prolif_50sim", label_col = "observed", direction = ">")
topo_res_prolif_mw_150sim = get_roc_stats(df = pred_topo_mw_bliss, pred_col = "synergy_prob_prolif_150sim", label_col = "observed", direction = ">")

# Plot ROCs
plot(x = topo_res_ss_ew_50sim$roc_stats$FPR, y = topo_res_ss_ew_50sim$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Ensemble-wise synergies (Bliss)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = topo_res_ss_ew_150sim$roc_stats$FPR, y = topo_res_ss_ew_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[2])
lines(x = topo_res_prolif_ew_50sim$roc_stats$FPR, y = topo_res_prolif_ew_50sim$roc_stats$TPR,
  lwd = 3, col = my_palette[3])
lines(x = topo_res_prolif_ew_150sim$roc_stats$FPR, y = topo_res_prolif_ew_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[4])
legend('bottomright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(topo_res_ss_ew_50sim$AUC, digits = 2), "Calibrated (50 sim)"),
    paste(round(topo_res_ss_ew_150sim$AUC, digits = 2), "Calibrated (150 sim)"),
    paste(round(topo_res_prolif_ew_50sim$AUC, digits = 2), "Proliferative (50 sim)"),
    paste(round(topo_res_prolif_ew_150sim$AUC, digits = 2), "Proliferative (150 sim)")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)

plot(x = topo_res_ss_mw_50sim$roc_stats$FPR, y = topo_res_ss_mw_50sim$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Model-wise synergies (Bliss)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = topo_res_ss_mw_150sim$roc_stats$FPR, y = topo_res_ss_mw_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[2])
lines(x = topo_res_prolif_mw_50sim$roc_stats$FPR, y = topo_res_prolif_mw_50sim$roc_stats$TPR,
  lwd = 3, col = my_palette[3])
lines(x = topo_res_prolif_mw_150sim$roc_stats$FPR, y = topo_res_prolif_mw_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[4])
legend('bottomright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(topo_res_ss_mw_50sim$AUC, digits = 2), "Calibrated (50 sim)"),
    paste(round(topo_res_ss_mw_150sim$AUC, digits = 2), "Calibrated (150 sim)"),
    paste(round(topo_res_prolif_mw_50sim$AUC, digits = 2), "Proliferative (50 sim)"),
    paste(round(topo_res_prolif_mw_150sim$AUC, digits = 2), "Proliferative (150 sim)")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)
```

## PR curves (Bliss) {-}

```{r PR curves Topology Mutations (CASCADE 2.0 - Bliss), fig.width=6, fig.height=6, dpi=300, cache=TRUE, fig.show='hold', out.width='50%'}
pr_topo_res_ss_ew_50sim = pr.curve(scores.class0 = pred_topo_ew_bliss %>% pull(ss_score_50sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topo_ew_bliss %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_topo_res_ss_ew_150sim = pr.curve(scores.class0 = pred_topo_ew_bliss %>% pull(ss_score_150sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topo_ew_bliss %>% pull(observed), curve = TRUE)
pr_topo_res_prolif_ew_50sim = pr.curve(scores.class0 = pred_topo_ew_bliss %>% pull(prolif_score_50sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topo_ew_bliss %>% pull(observed), curve = TRUE)
pr_topo_res_prolif_ew_150sim = pr.curve(scores.class0 = pred_topo_ew_bliss %>% pull(prolif_score_150sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topo_ew_bliss %>% pull(observed), curve = TRUE)

pr_topo_res_ss_mw_50sim = pr.curve(scores.class0 = pred_topo_mw_bliss %>% pull(synergy_prob_ss_50sim),
  weights.class0 = pred_topo_mw_bliss %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_topo_res_ss_mw_150sim = pr.curve(scores.class0 = pred_topo_mw_bliss %>% pull(synergy_prob_ss_150sim),
  weights.class0 = pred_topo_mw_bliss %>% pull(observed), curve = TRUE)
pr_topo_res_prolif_mw_50sim = pr.curve(scores.class0 = pred_topo_mw_bliss %>% pull(synergy_prob_prolif_50sim),
  weights.class0 = pred_topo_mw_bliss %>% pull(observed), curve = TRUE)
pr_topo_res_prolif_mw_150sim = pr.curve(scores.class0 = pred_topo_mw_bliss %>% pull(synergy_prob_prolif_150sim),
  weights.class0 = pred_topo_mw_bliss %>% pull(observed), curve = TRUE)

plot(pr_topo_res_ss_ew_50sim, main = 'PR curve, Ensemble-wise synergies (Bliss)',
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_topo_res_ss_ew_150sim, add = TRUE, color = my_palette[2])
plot(pr_topo_res_prolif_ew_50sim, add = TRUE, color = my_palette[3])
plot(pr_topo_res_prolif_ew_150sim, add = TRUE, color = my_palette[4])
legend('topright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(pr_topo_res_ss_ew_50sim$auc.davis.goadrich, digits = 3), "Calibrated (50 sim)"),
    paste(round(pr_topo_res_ss_ew_150sim$auc.davis.goadrich, digits = 3), "Calibrated (150 sim)"),
    paste(round(pr_topo_res_prolif_ew_50sim$auc.davis.goadrich, digits = 3), "Proliferative (50 sim)"),
    paste(round(pr_topo_res_prolif_ew_150sim$auc.davis.goadrich, digits = 3), "Proliferative (150 sim)")))
grid(lwd = 0.5)

plot(pr_topo_res_ss_mw_50sim, main = 'PR curve, Model-wise synergies (Bliss)',
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_topo_res_ss_mw_150sim, add = TRUE, color = my_palette[2])
plot(pr_topo_res_prolif_mw_50sim, add = TRUE, color = my_palette[3])
plot(pr_topo_res_prolif_mw_150sim, add = TRUE, color = my_palette[4])
legend('topright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(pr_topo_res_ss_mw_50sim$auc.davis.goadrich, digits = 3), "Calibrated (50 sim)"),
    paste(round(pr_topo_res_ss_mw_150sim$auc.davis.goadrich, digits = 3), "Calibrated (150 sim)"),
    paste(round(pr_topo_res_prolif_mw_50sim$auc.davis.goadrich, digits = 3), "Proliferative (50 sim)"),
    paste(round(pr_topo_res_prolif_mw_150sim$auc.davis.goadrich, digits = 3), "Proliferative (150 sim)")))
grid(lwd = 0.5)
```

## ROC-AUC sensitivity (Bliss) {-}

```{r ROC-AUC sensitivity Topology Mutations (Bliss - CASCADE 2.0), dpi=300, out.width='80%', fig.align='center', cache=TRUE}
betas = seq(from = -10, to = 10, by = 0.1)

auc_values_ew = sapply(betas, function(beta) {
  pred_topo_ew_bliss = pred_topo_ew_bliss %>% mutate(combined_score = ss_score_150sim + beta * prolif_score_150sim)
  res = get_roc_stats(df = pred_topo_ew_bliss, pred_col = "combined_score", label_col = "observed")
  auc_value = res$AUC
})

df_ew = as_tibble(cbind(betas, auc_values_ew))

ggline(data = df_ew, x = "betas", y = "auc_values_ew", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("$\\beta$"), ylab = "AUC (Area Under ROC Curve)",
  title = TeX("AUC sensitivity to $\\beta$ parameter: $calibrated + \\beta \\times random$"),
  color = my_palette[2], label = "betas", repel = TRUE,
  label.select = list(criteria = "`y` == max(`y`)"),
  font.label = list(size = 14)) + geom_vline(xintercept = 0) + grids()
```

## PR-AUC sensitivity (Bliss) {-}

```{r PR-AUC sensitivity Topology Mutations (Bliss - CASCADE 2.0), dpi=300, out.width='80%', fig.align='center', cache=TRUE}
# Ensemble-wise
betas = seq(from = -10, to = 10, by = 0.1)

auc_values_ew = sapply(betas, function(beta) {
  pred_topo_ew_bliss = pred_topo_ew_bliss %>% mutate(combined_score = ss_score_150sim + beta * prolif_score_150sim)
  res = pr.curve(scores.class0 = pred_topo_ew_bliss %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_topo_ew_bliss %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_ew = as_tibble(cbind(betas, auc_values_ew))

ggline(data = df_ew, x = "betas", y = "auc_values_ew", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("$\\beta$"), ylab = "AUC (Area Under PR Curve)",
  title = TeX("AUC sensitivity to $\\beta$ parameter: $calibrated + \\beta \\times random$"),
  color = my_palette[2], label = "betas", repel = TRUE,
  label.select = list(criteria = "`y` == max(`y`)"),
  font.label = list(size = 14)) + geom_vline(xintercept = 0) + grids()
```

## ROC and PRC for best beta (Bliss) {-}

```{r ROC and PRC for best beta - Topology Mutations (Bliss - CASCADE 2.0), fig.width=6, fig.height=6, dpi=300, cache=TRUE, fig.show='hold', out.width='50%'}
best_beta = -1
pred_topo_ew_bliss = pred_topo_ew_bliss %>% mutate(best_score = ss_score_150sim + best_beta * prolif_score_150sim)

roc_best_res = get_roc_stats(df = pred_topo_ew_bliss, pred_col = "best_score", label_col = "observed")
pr_best_res = pr.curve(scores.class0 = pred_topo_ew_bliss %>% pull(best_score) %>% (function(x) {-x}), 
    weights.class0 = pred_topo_ew_bliss %>% pull(observed), curve = TRUE, rand.compute = TRUE)

# Plot best ROC
plot(x = roc_best_res$roc_stats$FPR, y = roc_best_res$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = TeX('ROC curve (Ensemble-wise), $calibrated + \\beta \\times proliferative$'),
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
legend('bottomright', title = TeX('AUC ($\\beta$ = -1)'), col = my_palette[1], pch = 19,
  legend = paste(round(roc_best_res$AUC, digits = 2), 'Bliss (150 sim)'))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)

# Plot best PRC
plot(pr_best_res, main = TeX('PR curve (Ensemble-wise), $calibrated + \\beta \\times proliferative$'),
  auc.main = FALSE, color = my_palette[2], rand.plot = TRUE)
legend('topright', title = TeX('AUC ($\\beta$ = -1)'), col = my_palette[2], pch = 19,
  legend = paste(round(pr_best_res$auc.davis.goadrich, digits = 3), 'Bliss (150 sim)'))
grid(lwd = 0.5)
```

# CASCADE 2.0 Analysis (Topology and Link Operator Mutations) {-}

:::{.note}
We run `Gitsbe` simulations with $50$ topology mutations and $3000$ link operator mutations (bootstrap values, reduced to $10$ and $3$ respectively after models with stabla states have been found), both for $50$ and $150$ simulations and both **fitting to steady state** (calibrated models) and to a **proliferative phenotype** (so not random models but as close as it can get to that since we are discussing topology mutations).

Results were assessed via `Drabme`'s ensemble and modelwise results, using both the HSA and Bliss method.
:::

```{r Load Topology and Link Operator Mutation results}
# 'ss' => calibrated models, 'rand' => proliferative models (so not random but kind of!)
# 'ew' => ensemble-wise, 'mw' => modelwise

## HSA results ss
topolink_ss_hsa_ew_50sim_file = paste0("results/topo-and-link/cascade_2.0_ss_50sim_fixpoints_hsa_ensemblewise_synergies.tab")
topolink_ss_hsa_mw_50sim_file = paste0("results/topo-and-link/cascade_2.0_ss_50sim_fixpoints_hsa_modelwise_synergies.tab")
topolink_ss_hsa_ew_150sim_file = paste0("results/topo-and-link/cascade_2.0_ss_150sim_fixpoints_hsa_ensemblewise_synergies.tab")
topolink_ss_hsa_mw_150sim_file = paste0("results/topo-and-link/cascade_2.0_ss_150sim_fixpoints_hsa_modelwise_synergies.tab")

topolink_ss_hsa_ew_synergies_50sim = emba::get_synergy_scores(topolink_ss_hsa_ew_50sim_file)
topolink_ss_hsa_mw_synergies_50sim = emba::get_synergy_scores(topolink_ss_hsa_mw_50sim_file, file_type = "modelwise")
topolink_ss_hsa_ew_synergies_150sim = emba::get_synergy_scores(topolink_ss_hsa_ew_150sim_file)
topolink_ss_hsa_mw_synergies_150sim = emba::get_synergy_scores(topolink_ss_hsa_mw_150sim_file, file_type = "modelwise")

## HSA results rand
topolink_prolif_hsa_ew_50sim_file = paste0("results/topo-and-link/cascade_2.0_rand_50sim_fixpoints_hsa_ensemblewise_synergies.tab")
topolink_prolif_hsa_mw_50sim_file = paste0("results/topo-and-link/cascade_2.0_rand_50sim_fixpoints_hsa_modelwise_synergies.tab")
topolink_prolif_hsa_ew_150sim_file = paste0("results/topo-and-link/cascade_2.0_rand_150sim_fixpoints_hsa_ensemblewise_synergies.tab")
topolink_prolif_hsa_mw_150sim_file = paste0("results/topo-and-link/cascade_2.0_rand_150sim_fixpoints_hsa_modelwise_synergies.tab")

topolink_prolif_hsa_ew_synergies_50sim = emba::get_synergy_scores(topolink_prolif_hsa_ew_50sim_file)
topolink_prolif_hsa_mw_synergies_50sim = emba::get_synergy_scores(topolink_prolif_hsa_mw_50sim_file, file_type = "modelwise")
topolink_prolif_hsa_ew_synergies_150sim = emba::get_synergy_scores(topolink_prolif_hsa_ew_150sim_file)
topolink_prolif_hsa_mw_synergies_150sim = emba::get_synergy_scores(topolink_prolif_hsa_mw_150sim_file, file_type = "modelwise")

## Bliss results ss
topolink_ss_bliss_ew_50sim_file = paste0("results/topo-and-link/cascade_2.0_ss_50sim_fixpoints_bliss_ensemblewise_synergies.tab")
topolink_ss_bliss_mw_50sim_file = paste0("results/topo-and-link/cascade_2.0_ss_50sim_fixpoints_bliss_modelwise_synergies.tab")
topolink_ss_bliss_ew_150sim_file = paste0("results/topo-and-link/cascade_2.0_ss_150sim_fixpoints_bliss_ensemblewise_synergies.tab")
topolink_ss_bliss_mw_150sim_file = paste0("results/topo-and-link/cascade_2.0_ss_150sim_fixpoints_bliss_modelwise_synergies.tab")

topolink_ss_bliss_ew_synergies_50sim = emba::get_synergy_scores(topolink_ss_bliss_ew_50sim_file)
topolink_ss_bliss_mw_synergies_50sim = emba::get_synergy_scores(topolink_ss_bliss_mw_50sim_file, file_type = "modelwise")
topolink_ss_bliss_ew_synergies_150sim = emba::get_synergy_scores(topolink_ss_bliss_ew_150sim_file)
topolink_ss_bliss_mw_synergies_150sim = emba::get_synergy_scores(topolink_ss_bliss_mw_150sim_file, file_type = "modelwise")

## Bliss results rand
topolink_prolif_bliss_ew_50sim_file = paste0("results/topo-and-link/cascade_2.0_rand_50sim_fixpoints_bliss_ensemblewise_synergies.tab")
topolink_prolif_bliss_mw_50sim_file = paste0("results/topo-and-link/cascade_2.0_rand_50sim_fixpoints_bliss_modelwise_synergies.tab")
topolink_prolif_bliss_ew_150sim_file = paste0("results/topo-and-link/cascade_2.0_rand_150sim_fixpoints_bliss_ensemblewise_synergies.tab")
topolink_prolif_bliss_mw_150sim_file = paste0("results/topo-and-link/cascade_2.0_rand_150sim_fixpoints_bliss_modelwise_synergies.tab")

topolink_prolif_bliss_ew_synergies_50sim = emba::get_synergy_scores(topolink_prolif_bliss_ew_50sim_file)
topolink_prolif_bliss_mw_synergies_50sim = emba::get_synergy_scores(topolink_prolif_bliss_mw_50sim_file, file_type = "modelwise")
topolink_prolif_bliss_ew_synergies_150sim = emba::get_synergy_scores(topolink_prolif_bliss_ew_150sim_file)
topolink_prolif_bliss_mw_synergies_150sim = emba::get_synergy_scores(topolink_prolif_bliss_mw_150sim_file, file_type = "modelwise")

# calculate probability of synergy in the modelwise results
topolink_ss_hsa_mw_synergies_50sim = topolink_ss_hsa_mw_synergies_50sim %>% 
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topolink_ss_hsa_mw_synergies_150sim = topolink_ss_hsa_mw_synergies_150sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topolink_prolif_hsa_mw_synergies_50sim = topolink_prolif_hsa_mw_synergies_50sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topolink_prolif_hsa_mw_synergies_150sim = topolink_prolif_hsa_mw_synergies_150sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topolink_ss_bliss_mw_synergies_50sim = topolink_ss_bliss_mw_synergies_50sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topolink_ss_bliss_mw_synergies_150sim = topolink_ss_bliss_mw_synergies_150sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topolink_prolif_bliss_mw_synergies_50sim = topolink_prolif_bliss_mw_synergies_50sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))
topolink_prolif_bliss_mw_synergies_150sim = topolink_prolif_bliss_mw_synergies_150sim %>%
  mutate(synergy_prob_ss = synergies/(synergies + `non-synergies`))

# Tidy the data
pred_topolink_ew_hsa = bind_cols(
  topolink_ss_hsa_ew_synergies_50sim %>% rename(ss_score_50sim = score),
  topolink_ss_hsa_ew_synergies_150sim %>% select(score) %>% rename(ss_score_150sim = score),
  topolink_prolif_hsa_ew_synergies_50sim %>% select(score) %>% rename(prolif_score_50sim = score),
  topolink_prolif_hsa_ew_synergies_150sim %>% select(score) %>% rename(prolif_score_150sim = score),
  as_tibble_col(observed, column_name = "observed"))

pred_topolink_mw_hsa = bind_cols(
  topolink_ss_hsa_mw_synergies_50sim %>% select(perturbation, synergy_prob_ss) %>% rename(synergy_prob_ss_50sim = synergy_prob_ss),
  topolink_ss_hsa_mw_synergies_150sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_ss_150sim = synergy_prob_ss),
  topolink_prolif_hsa_mw_synergies_50sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_prolif_50sim = synergy_prob_ss),
  topolink_prolif_hsa_mw_synergies_150sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_prolif_150sim = synergy_prob_ss),
  as_tibble_col(observed, column_name = "observed"))

pred_topolink_ew_bliss = bind_cols(
  topolink_ss_bliss_ew_synergies_50sim %>% rename(ss_score_50sim = score),
  topolink_ss_bliss_ew_synergies_150sim %>% select(score) %>% rename(ss_score_150sim = score),
  topolink_prolif_bliss_ew_synergies_50sim %>% select(score) %>% rename(prolif_score_50sim = score),
  topolink_prolif_bliss_ew_synergies_150sim %>% select(score) %>% rename(prolif_score_150sim = score),
  as_tibble_col(observed, column_name = "observed"))

pred_topolink_mw_bliss = bind_cols(
  topolink_ss_bliss_mw_synergies_50sim %>% select(perturbation, synergy_prob_ss) %>% rename(synergy_prob_ss_50sim = synergy_prob_ss),
  topolink_ss_bliss_mw_synergies_150sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_ss_150sim = synergy_prob_ss),
  topolink_prolif_bliss_mw_synergies_50sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_prolif_50sim = synergy_prob_ss),
  topolink_prolif_bliss_mw_synergies_150sim %>% select(synergy_prob_ss) %>% rename(synergy_prob_prolif_150sim = synergy_prob_ss),
  as_tibble_col(observed, column_name = "observed"))
```

### ROC curves (HSA) {-}

```{r ROC curves Topology and Link Operator Mutations (CASCADE 2.0 - HSA), fig.width=6, fig.height=6, dpi=300, cache=TRUE, fig.show='hold', out.width='50%'}
topolink_res_ss_ew_50sim = get_roc_stats(df = pred_topolink_ew_hsa, pred_col = "ss_score_50sim", label_col = "observed")
topolink_res_ss_ew_150sim = get_roc_stats(df = pred_topolink_ew_hsa, pred_col = "ss_score_150sim", label_col = "observed")
topolink_res_prolif_ew_50sim = get_roc_stats(df = pred_topolink_ew_hsa, pred_col = "prolif_score_50sim", label_col = "observed")
topolink_res_prolif_ew_150sim = get_roc_stats(df = pred_topolink_ew_hsa, pred_col = "prolif_score_150sim", label_col = "observed")

topolink_res_ss_mw_50sim = get_roc_stats(df = pred_topolink_mw_hsa, pred_col = "synergy_prob_ss_50sim", label_col = "observed", direction = ">")
topolink_res_ss_mw_150sim = get_roc_stats(df = pred_topolink_mw_hsa, pred_col = "synergy_prob_ss_150sim", label_col = "observed", direction = ">")
topolink_res_prolif_mw_50sim = get_roc_stats(df = pred_topolink_mw_hsa, pred_col = "synergy_prob_prolif_50sim", label_col = "observed", direction = ">")
topolink_res_prolif_mw_150sim = get_roc_stats(df = pred_topolink_mw_hsa, pred_col = "synergy_prob_prolif_150sim", label_col = "observed", direction = ">")

# Plot ROCs
plot(x = topolink_res_ss_ew_50sim$roc_stats$FPR, y = topolink_res_ss_ew_50sim$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Ensemble-wise synergies (HSA)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = topolink_res_ss_ew_150sim$roc_stats$FPR, y = topolink_res_ss_ew_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[2])
lines(x = topolink_res_prolif_ew_50sim$roc_stats$FPR, y = topolink_res_prolif_ew_50sim$roc_stats$TPR,
  lwd = 3, col = my_palette[3])
lines(x = topolink_res_prolif_ew_150sim$roc_stats$FPR, y = topolink_res_prolif_ew_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[4])
legend('bottomright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(topolink_res_ss_ew_50sim$AUC, digits = 2), "Calibrated (50 sim)"),
    paste(round(topolink_res_ss_ew_150sim$AUC, digits = 2), "Calibrated (150 sim)"),
    paste(round(topolink_res_prolif_ew_50sim$AUC, digits = 2), "Proliferative (50 sim)"),
    paste(round(topolink_res_prolif_ew_150sim$AUC, digits = 2), "Proliferative (150 sim)")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)

plot(x = topolink_res_ss_mw_50sim$roc_stats$FPR, y = topolink_res_ss_mw_50sim$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Model-wise synergies (HSA)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = topolink_res_ss_mw_150sim$roc_stats$FPR, y = topolink_res_ss_mw_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[2])
lines(x = topolink_res_prolif_mw_50sim$roc_stats$FPR, y = topolink_res_prolif_mw_50sim$roc_stats$TPR,
  lwd = 3, col = my_palette[3])
lines(x = topolink_res_prolif_mw_150sim$roc_stats$FPR, y = topolink_res_prolif_mw_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[4])
legend('bottomright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(topolink_res_ss_mw_50sim$AUC, digits = 2), "Calibrated (50 sim)"),
    paste(round(topolink_res_ss_mw_150sim$AUC, digits = 2), "Calibrated (150 sim)"),
    paste(round(topolink_res_prolif_mw_50sim$AUC, digits = 2), "Proliferative (50 sim)"),
    paste(round(topolink_res_prolif_mw_150sim$AUC, digits = 2), "Proliferative (150 sim)")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)
```

### PR curves (HSA) {-}

```{r PR curves Topology and Link Mutations (CASCADE 2.0 - HSA), fig.width=6, fig.height=6, dpi=300, cache=TRUE, fig.show='hold', out.width='50%'}
pr_topolink_res_ss_ew_50sim = pr.curve(scores.class0 = pred_topolink_ew_hsa %>% pull(ss_score_50sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topolink_ew_hsa %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_topolink_res_ss_ew_150sim = pr.curve(scores.class0 = pred_topolink_ew_hsa %>% pull(ss_score_150sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topolink_ew_hsa %>% pull(observed), curve = TRUE)
pr_topolink_res_prolif_ew_50sim = pr.curve(scores.class0 = pred_topolink_ew_hsa %>% pull(prolif_score_50sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topolink_ew_hsa %>% pull(observed), curve = TRUE)
pr_topolink_res_prolif_ew_150sim = pr.curve(scores.class0 = pred_topolink_ew_hsa %>% pull(prolif_score_150sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topolink_ew_hsa %>% pull(observed), curve = TRUE)

pr_topolink_res_ss_mw_50sim = pr.curve(scores.class0 = pred_topolink_mw_hsa %>% pull(synergy_prob_ss_50sim),
  weights.class0 = pred_topolink_mw_hsa %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_topolink_res_ss_mw_150sim = pr.curve(scores.class0 = pred_topolink_mw_hsa %>% pull(synergy_prob_ss_150sim),
  weights.class0 = pred_topolink_mw_hsa %>% pull(observed), curve = TRUE)
pr_topolink_res_prolif_mw_50sim = pr.curve(scores.class0 = pred_topolink_mw_hsa %>% pull(synergy_prob_prolif_50sim),
  weights.class0 = pred_topolink_mw_hsa %>% pull(observed), curve = TRUE)
pr_topolink_res_prolif_mw_150sim = pr.curve(scores.class0 = pred_topolink_mw_hsa %>% pull(synergy_prob_prolif_150sim),
  weights.class0 = pred_topolink_mw_hsa %>% pull(observed), curve = TRUE)

plot(pr_topolink_res_ss_ew_50sim, main = 'PR curve, Ensemble-wise synergies (HSA)',
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_topolink_res_ss_ew_150sim, add = TRUE, color = my_palette[2])
plot(pr_topolink_res_prolif_ew_50sim, add = TRUE, color = my_palette[3])
plot(pr_topolink_res_prolif_ew_150sim, add = TRUE, color = my_palette[4])
legend('topright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(pr_topolink_res_ss_ew_50sim$auc.davis.goadrich, digits = 3), "Calibrated (50 sim)"),
    paste(round(pr_topolink_res_ss_ew_150sim$auc.davis.goadrich, digits = 3), "Calibrated (150 sim)"),
    paste(round(pr_topolink_res_prolif_ew_50sim$auc.davis.goadrich, digits = 3), "Proliferative (50 sim)"),
    paste(round(pr_topolink_res_prolif_ew_150sim$auc.davis.goadrich, digits = 3), "Proliferative (150 sim)")))
grid(lwd = 0.5)

plot(pr_topolink_res_ss_mw_50sim, main = 'PR curve, Model-wise synergies (HSA)',
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_topolink_res_ss_mw_150sim, add = TRUE, color = my_palette[2])
plot(pr_topolink_res_prolif_mw_50sim, add = TRUE, color = my_palette[3])
plot(pr_topolink_res_prolif_mw_150sim, add = TRUE, color = my_palette[4])
legend('topright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(pr_topolink_res_ss_mw_50sim$auc.davis.goadrich, digits = 3), "Calibrated (50 sim)"),
    paste(round(pr_topolink_res_ss_mw_150sim$auc.davis.goadrich, digits = 3), "Calibrated (150 sim)"),
    paste(round(pr_topolink_res_prolif_mw_50sim$auc.davis.goadrich, digits = 3), "Proliferative (50 sim)"),
    paste(round(pr_topolink_res_prolif_mw_150sim$auc.davis.goadrich, digits = 3), "Proliferative (150 sim)")))
grid(lwd = 0.5)
```

### ROC-AUC sensitivity (HSA) {-}

Combine the $150$ simulation results (calibrated + proliferative)
```{r ROC-AUC sensitivity Topology and Link Mutations (HSA - CASCADE 2.0), dpi=300, out.width='80%', fig.align='center', cache=TRUE}
# Ensemble-wise
betas = seq(from = -10, to = 10, by = 0.1)

auc_values_ew = sapply(betas, function(beta) {
  pred_topolink_ew_hsa = pred_topolink_ew_hsa %>% mutate(combined_score = ss_score_150sim + beta * prolif_score_150sim)
  res = get_roc_stats(df = pred_topolink_ew_hsa, pred_col = "combined_score", label_col = "observed")
  auc_value = res$AUC
})

df_ew = as_tibble(cbind(betas, auc_values_ew))

ggline(data = df_ew, x = "betas", y = "auc_values_ew", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("$\\beta$"), ylab = "AUC (Area Under ROC Curve)",
  title = TeX("AUC sensitivity to $\\beta$ parameter: $calibrated + \\beta \\times random$"),
  color = my_palette[2], label = "betas", label.rectangle = TRUE, repel = TRUE,
  label.select = list(criteria = "`y` == max(`y`)"),
  font.label = list(size = 14)) + geom_vline(xintercept = 0) + grids()

# Model-wise
weights = seq(from = 0, to = 1, by = 0.05)

auc_values_mw = sapply(weights, function(w) {
  pred_topolink_mw_hsa = pred_topolink_mw_hsa %>% 
    mutate(weighted_prob = (1 - w) * pred_topolink_mw_hsa$synergy_prob_ss_150sim + w * pred_topolink_mw_hsa$synergy_prob_prolif_150sim)
  res = get_roc_stats(df = pred_topolink_mw_hsa, pred_col = "weighted_prob", label_col = "observed", direction = ">")
  auc_value = res$AUC
})

df_mw = as_tibble(cbind(weights, auc_values_mw))

ggline(data = df_mw, x = "weights", y = "auc_values_mw", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("weight $w$"), ylab = "AUC (Area Under ROC Curve)",
  title = TeX("AUC sensitivity to weighted average score: $(1-w) \\times prob_{ss} + w \\times prob_{rand}$"),
  color = my_palette[3]) + grids()
```

### PR-AUC sensitivity (HSA) {-}

Combine the $150$ simulation results (calibrated + proliferative)
```{r PR-AUC sensitivity Topology and Link Mutations (HSA - CASCADE 2.0), dpi=300, out.width='80%', fig.align='center', cache=TRUE}
# Ensemble-wise
betas = seq(from = -10, to = 10, by = 0.1)

auc_values_ew = sapply(betas, function(beta) {
  pred_topolink_ew_hsa = pred_topolink_ew_hsa %>% mutate(combined_score = ss_score_150sim + beta * prolif_score_150sim)
  res = pr.curve(scores.class0 = pred_topolink_ew_hsa %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_topolink_ew_hsa %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_ew = as_tibble(cbind(betas, auc_values_ew))

ggline(data = df_ew, x = "betas", y = "auc_values_ew", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("$\\beta$"), ylab = "AUC (Area Under PR Curve)",
  title = TeX("AUC sensitivity to $\\beta$ parameter: $calibrated + \\beta \\times random$"),
  color = my_palette[2], label = "betas", label.rectangle = TRUE, repel = TRUE,
  label.select = list(criteria = "`y` == max(`y`)"),
  font.label = list(size = 14)) + geom_vline(xintercept = 0) + grids()

# Model-wise
weights = seq(from = 0, to = 1, by = 0.05)

auc_values_mw = sapply(weights, function(w) {
  pred_topolink_mw_hsa = pred_topolink_mw_hsa %>% 
    mutate(weighted_prob = (1 - w) * pred_topolink_mw_hsa$synergy_prob_ss_150sim + w * pred_topolink_mw_hsa$synergy_prob_prolif_150sim)
  res = pr.curve(scores.class0 = pred_topolink_mw_hsa %>% pull(weighted_prob), 
    weights.class0 = pred_topolink_mw_hsa %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_mw = as_tibble(cbind(weights, auc_values_mw))

ggline(data = df_mw, x = "weights", y = "auc_values_mw", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("weight $w$"), ylab = "AUC (Area Under PR Curve)",
  title = TeX("AUC sensitivity to weighted average score: $(1-w) \\times prob_{ss} + w \\times prob_{rand}$"),
  color = my_palette[3]) + grids()
```

### ROC curves (Bliss) {-}

```{r ROC curves Topology and Link Mutations (CASCADE 2.0 - Bliss), fig.width=6, fig.height=6, dpi=300, cache=TRUE, fig.show='hold', out.width='50%'}
topolink_res_ss_ew_50sim = get_roc_stats(df = pred_topolink_ew_bliss, pred_col = "ss_score_50sim", label_col = "observed")
topolink_res_ss_ew_150sim = get_roc_stats(df = pred_topolink_ew_bliss, pred_col = "ss_score_150sim", label_col = "observed")
topolink_res_prolif_ew_50sim = get_roc_stats(df = pred_topolink_ew_bliss, pred_col = "prolif_score_50sim", label_col = "observed")
topolink_res_prolif_ew_150sim = get_roc_stats(df = pred_topolink_ew_bliss, pred_col = "prolif_score_150sim", label_col = "observed")

topolink_res_ss_mw_50sim = get_roc_stats(df = pred_topolink_mw_bliss, pred_col = "synergy_prob_ss_50sim", label_col = "observed", direction = ">")
topolink_res_ss_mw_150sim = get_roc_stats(df = pred_topolink_mw_bliss, pred_col = "synergy_prob_ss_150sim", label_col = "observed", direction = ">")
topolink_res_prolif_mw_50sim = get_roc_stats(df = pred_topolink_mw_bliss, pred_col = "synergy_prob_prolif_50sim", label_col = "observed", direction = ">")
topolink_res_prolif_mw_150sim = get_roc_stats(df = pred_topolink_mw_bliss, pred_col = "synergy_prob_prolif_150sim", label_col = "observed", direction = ">")

# Plot ROCs
plot(x = topolink_res_ss_ew_50sim$roc_stats$FPR, y = topolink_res_ss_ew_50sim$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Ensemble-wise synergies (Bliss)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = topolink_res_ss_ew_150sim$roc_stats$FPR, y = topolink_res_ss_ew_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[2])
lines(x = topolink_res_prolif_ew_50sim$roc_stats$FPR, y = topolink_res_prolif_ew_50sim$roc_stats$TPR,
  lwd = 3, col = my_palette[3])
lines(x = topolink_res_prolif_ew_150sim$roc_stats$FPR, y = topolink_res_prolif_ew_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[4])
legend('bottomright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(topolink_res_ss_ew_50sim$AUC, digits = 2), "Calibrated (50 sim)"),
    paste(round(topolink_res_ss_ew_150sim$AUC, digits = 2), "Calibrated (150 sim)"),
    paste(round(topolink_res_prolif_ew_50sim$AUC, digits = 2), "Proliferative (50 sim)"),
    paste(round(topolink_res_prolif_ew_150sim$AUC, digits = 2), "Proliferative (150 sim)")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)

plot(x = topolink_res_ss_mw_50sim$roc_stats$FPR, y = topolink_res_ss_mw_50sim$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = 'ROC curve, Model-wise synergies (Bliss)',
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = topolink_res_ss_mw_150sim$roc_stats$FPR, y = topolink_res_ss_mw_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[2])
lines(x = topolink_res_prolif_mw_50sim$roc_stats$FPR, y = topolink_res_prolif_mw_50sim$roc_stats$TPR,
  lwd = 3, col = my_palette[3])
lines(x = topolink_res_prolif_mw_150sim$roc_stats$FPR, y = topolink_res_prolif_mw_150sim$roc_stats$TPR,
  lwd = 3, col = my_palette[4])
legend('bottomright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(topolink_res_ss_mw_50sim$AUC, digits = 2), "Calibrated (50 sim)"),
    paste(round(topolink_res_ss_mw_150sim$AUC, digits = 2), "Calibrated (150 sim)"),
    paste(round(topolink_res_prolif_mw_50sim$AUC, digits = 2), "Proliferative (50 sim)"),
    paste(round(topolink_res_prolif_mw_150sim$AUC, digits = 2), "Proliferative (150 sim)")))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)
```

### PR curves (Bliss) {-}

```{r PR curves Topology and Link Mutations (CASCADE 2.0 - Bliss), fig.width=6, fig.height=6, dpi=300, cache=TRUE, fig.show='hold', out.width='50%'}
pr_topolink_res_ss_ew_50sim = pr.curve(scores.class0 = pred_topolink_ew_bliss %>% pull(ss_score_50sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topolink_ew_bliss %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_topolink_res_ss_ew_150sim = pr.curve(scores.class0 = pred_topolink_ew_bliss %>% pull(ss_score_150sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topolink_ew_bliss %>% pull(observed), curve = TRUE)
pr_topolink_res_prolif_ew_50sim = pr.curve(scores.class0 = pred_topolink_ew_bliss %>% pull(prolif_score_50sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topolink_ew_bliss %>% pull(observed), curve = TRUE)
pr_topolink_res_prolif_ew_150sim = pr.curve(scores.class0 = pred_topolink_ew_bliss %>% pull(prolif_score_150sim) %>% (function(x) {-x}), 
  weights.class0 = pred_topolink_ew_bliss %>% pull(observed), curve = TRUE)

pr_topolink_res_ss_mw_50sim = pr.curve(scores.class0 = pred_topolink_mw_bliss %>% pull(synergy_prob_ss_50sim),
  weights.class0 = pred_topolink_mw_bliss %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_topolink_res_ss_mw_150sim = pr.curve(scores.class0 = pred_topolink_mw_bliss %>% pull(synergy_prob_ss_150sim),
  weights.class0 = pred_topolink_mw_bliss %>% pull(observed), curve = TRUE)
pr_topolink_res_prolif_mw_50sim = pr.curve(scores.class0 = pred_topolink_mw_bliss %>% pull(synergy_prob_prolif_50sim),
  weights.class0 = pred_topolink_mw_bliss %>% pull(observed), curve = TRUE)
pr_topolink_res_prolif_mw_150sim = pr.curve(scores.class0 = pred_topolink_mw_bliss %>% pull(synergy_prob_prolif_150sim),
  weights.class0 = pred_topolink_mw_bliss %>% pull(observed), curve = TRUE)

plot(pr_topolink_res_ss_ew_50sim, main = 'PR curve, Ensemble-wise synergies (Bliss)',
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_topolink_res_ss_ew_150sim, add = TRUE, color = my_palette[2])
plot(pr_topolink_res_prolif_ew_50sim, add = TRUE, color = my_palette[3])
plot(pr_topolink_res_prolif_ew_150sim, add = TRUE, color = my_palette[4])
legend('topright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(pr_topolink_res_ss_ew_50sim$auc.davis.goadrich, digits = 3), "Calibrated (50 sim)"),
    paste(round(pr_topolink_res_ss_ew_150sim$auc.davis.goadrich, digits = 3), "Calibrated (150 sim)"),
    paste(round(pr_topolink_res_prolif_ew_50sim$auc.davis.goadrich, digits = 3), "Proliferative (50 sim)"),
    paste(round(pr_topolink_res_prolif_ew_150sim$auc.davis.goadrich, digits = 3), "Proliferative (150 sim)")))
grid(lwd = 0.5)

plot(pr_topolink_res_ss_mw_50sim, main = 'PR curve, Model-wise synergies (Bliss)',
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_topolink_res_ss_mw_150sim, add = TRUE, color = my_palette[2])
plot(pr_topolink_res_prolif_mw_50sim, add = TRUE, color = my_palette[3])
plot(pr_topolink_res_prolif_mw_150sim, add = TRUE, color = my_palette[4])
legend('topright', title = 'AUC', col = my_palette[1:4], pch = 19,
  legend = c(paste(round(pr_topolink_res_ss_mw_50sim$auc.davis.goadrich, digits = 3), "Calibrated (50 sim)"),
    paste(round(pr_topolink_res_ss_mw_150sim$auc.davis.goadrich, digits = 3), "Calibrated (150 sim)"),
    paste(round(pr_topolink_res_prolif_mw_50sim$auc.davis.goadrich, digits = 3), "Proliferative (50 sim)"),
    paste(round(pr_topolink_res_prolif_mw_150sim$auc.davis.goadrich, digits = 3), "Proliferative (150 sim)")))
grid(lwd = 0.5)
```

### ROC-AUC sensitivity (Bliss) {-}

Combine the $150$ simulation results (calibrated + proliferative)
```{r ROC-AUC sensitivity Topology and Link Mutations (Bliss - CASCADE 2.0), dpi=300, out.width='80%', fig.align='center', cache=TRUE}
# Ensemble-wise
betas = seq(from = -10, to = 10, by = 0.1)

auc_values_ew = sapply(betas, function(beta) {
  pred_topolink_ew_bliss = pred_topolink_ew_bliss %>% mutate(combined_score = ss_score_150sim + beta * prolif_score_150sim)
  res = get_roc_stats(df = pred_topolink_ew_bliss, pred_col = "combined_score", label_col = "observed")
  auc_value = res$AUC
})

df_ew = as_tibble(cbind(betas, auc_values_ew))

ggline(data = df_ew, x = "betas", y = "auc_values_ew", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("$\\beta$"), ylab = "AUC (Area Under ROC Curve)",
  title = TeX("AUC sensitivity to $\\beta$ parameter: $calibrated + \\beta \\times random$"),
  color = my_palette[2], label = "betas", repel = TRUE,
  label.select = list(criteria = "`y` == max(`y`)"),
  font.label = list(size = 14)) + geom_vline(xintercept = 0) + grids()

# Model-wise
weights = seq(from = 0, to = 1, by = 0.05)

auc_values_mw = sapply(weights, function(w) {
  pred_topolink_mw_bliss = pred_topolink_mw_bliss %>% 
    mutate(weighted_prob = (1 - w) * pred_topolink_mw_bliss$synergy_prob_ss_150sim + w * pred_topolink_mw_bliss$synergy_prob_prolif_150sim)
  res = get_roc_stats(df = pred_topolink_mw_bliss, pred_col = "weighted_prob", label_col = "observed", direction = ">")
  auc_value = res$AUC
})

df_mw = as_tibble(cbind(weights, auc_values_mw))

ggline(data = df_mw, x = "weights", y = "auc_values_mw", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("weight $w$"), ylab = "AUC (Area Under ROC Curve)",
  title = TeX("AUC sensitivity to weighted average score: $(1-w) \\times prob_{ss} + w \\times prob_{rand}$"),
  color = my_palette[3]) + grids()
```

### PR-AUC sensitivity (Bliss) {-}

Combine the $150$ simulation results (calibrated + proliferative)
```{r PR-AUC sensitivity Topology and Link Mutations (Bliss - CASCADE 2.0), dpi=300, out.width='80%', fig.align='center', cache=TRUE}
# Ensemble-wise
betas = seq(from = -10, to = 10, by = 0.1)

auc_values_ew = sapply(betas, function(beta) {
  pred_topolink_ew_bliss = pred_topolink_ew_bliss %>% mutate(combined_score = ss_score_150sim + beta * prolif_score_150sim)
  res = pr.curve(scores.class0 = pred_topolink_ew_bliss %>% pull(combined_score) %>% (function(x) {-x}), 
    weights.class0 = pred_topolink_ew_bliss %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_ew = as_tibble(cbind(betas, auc_values_ew))

ggline(data = df_ew, x = "betas", y = "auc_values_ew", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("$\\beta$"), ylab = "AUC (Area Under PR Curve)",
  title = TeX("AUC sensitivity to $\\beta$ parameter: $calibrated + \\beta \\times random$"),
  color = my_palette[2], label = "betas", repel = TRUE,
  label.select = list(criteria = "`y` == max(`y`)"),
  font.label = list(size = 14)) + geom_vline(xintercept = 0) + grids()

# Model-wise
weights = seq(from = 0, to = 1, by = 0.05)

auc_values_mw = sapply(weights, function(w) {
  pred_topolink_mw_bliss = pred_topolink_mw_bliss %>% 
    mutate(weighted_prob = (1 - w) * pred_topolink_mw_bliss$synergy_prob_ss_150sim + w * pred_topolink_mw_bliss$synergy_prob_prolif_150sim)
  res = pr.curve(scores.class0 = pred_topolink_mw_bliss %>% pull(weighted_prob), 
    weights.class0 = pred_topolink_mw_bliss %>% pull(observed))
  auc_value = res$auc.davis.goadrich
})

df_mw = as_tibble(cbind(weights, auc_values_mw))

ggline(data = df_mw, x = "weights", y = "auc_values_mw", numeric.x.axis = TRUE,
  plot_type = "l", xlab = TeX("weight $w$"), ylab = "AUC (Area Under PR Curve)",
  title = TeX("AUC sensitivity to weighted average score: $(1-w) \\times prob_{ss} + w \\times prob_{rand}$"),
  color = my_palette[3]) + grids()
```

### ROC and PRC for best beta (HSA & Bliss) {-}

Use the best betas from the PR-AUC sensitivity graphs (separately for Bliss and HSA)!

```{r ROC and PRC for best beta - Topology and Link Mutations (HSA and Bliss - CASCADE 2.0), fig.width=6, fig.height=6, dpi=300, cache=TRUE, fig.show='hold', out.width='50%'}
best_beta_hsa = -1
best_best_bliss = -1.3
pred_topolink_ew_hsa = pred_topolink_ew_hsa %>% mutate(best_score = ss_score_150sim + best_beta_hsa * prolif_score_150sim)
pred_topolink_ew_bliss = pred_topolink_ew_bliss %>% mutate(best_score = ss_score_150sim + best_best_bliss * prolif_score_150sim)

roc_best_res_hsa = get_roc_stats(df = pred_topolink_ew_hsa, pred_col = "best_score", label_col = "observed")
roc_best_res_bliss = get_roc_stats(df = pred_topolink_ew_bliss, pred_col = "best_score", label_col = "observed")

pr_best_res_hsa = pr.curve(scores.class0 = pred_topolink_ew_hsa %>% pull(best_score) %>% (function(x) {-x}), 
    weights.class0 = pred_topolink_ew_hsa %>% pull(observed), curve = TRUE, rand.compute = TRUE)
pr_best_res_bliss = pr.curve(scores.class0 = pred_topolink_ew_bliss %>% pull(best_score) %>% (function(x) {-x}), 
    weights.class0 = pred_topolink_ew_bliss %>% pull(observed), curve = TRUE)

# Plot best ROCs
plot(x = roc_best_res_hsa$roc_stats$FPR, y = roc_best_res_hsa$roc_stats$TPR,
  type = 'l', lwd = 3, col = my_palette[1], main = TeX('ROC curve (Ensemble-wise), $calibrated + \\beta \\times proliferative$'),
  xlab = 'False Positive Rate (FPR)', ylab = 'True Positive Rate (TPR)')
lines(x = roc_best_res_bliss$roc_stats$FPR, y = roc_best_res_bliss$roc_stats$TPR,
  lwd = 3, col = my_palette[3])
legend('bottomright', title = TeX('AUC ($\\beta_{HSA}$ = -1, $\\beta_{Bliss}$ = -1.3)'), 
  col = c(my_palette[1], my_palette[3]), pch = 19,
  legend = c(paste(round(roc_best_res_hsa$AUC, digits = 2), 'HSA (150 sim)'),
    paste(round(roc_best_res_bliss$AUC, digits = 2), 'Bliss (150 sim)')))
grid(lwd = 0.5)
abline(a = 0, b = 1, col = 'lightgrey', lty = 'dotdash', lwd = 1.2)

# Plot best PRCs
plot(pr_best_res_hsa, main = TeX('PR curve (Ensemble-wise), $calibrated + \\beta \\times proliferative$'),
  auc.main = FALSE, color = my_palette[1], rand.plot = TRUE)
plot(pr_best_res_bliss, add = TRUE, color = my_palette[3])
legend('topright', title = TeX('AUC ($\\beta_{HSA}$ = -1, $\\beta_{Bliss}$ = -1.3)'), col = c(my_palette[1], my_palette[3]), pch = 19,
  legend = c(paste(round(pr_best_res_hsa$auc.davis.goadrich, digits = 2), 'HSA (150 sim)'),
    paste(round(pr_best_res_bliss$auc.davis.goadrich, digits = 2), 'Bliss (150 sim)')))
grid(lwd = 0.5)
```

# Reproduce simulation results {-}

## ROC curves {-}

- Install `druglogics-synergy`
- Run the script `run_druglogics_synergy.sh` in the above repo using the configuration settings: 
  - `simulations: 50`
  - `attractor_tool: biolqm_stable_states`
  - `synergy_method: hsa` (also rerun the script chaning the synergy method `bliss`)

Thus you will get a directory per simulation and inside will be several result files.
To generate the ROC curves, we use the ensemble-wise and model-wise synergies found in each respective simulation.

## Random model results {-}

Use the `druglogics-synergy` module, version `1.2.0`: `git checkout v1.2.0` and the [abmlog](https://github.com/druglogics/abmlog) module, version `1.5.0`: `git checkout v1.5.0`.

The CASCADE 1.0 and 2.0 `.sif` network files can be found at the directories `ags_cascade_1.0` and `ags_cascade_2.0` on the
`druglogics-synergy` repository.
Copy these two network files inside the `test` dir of the `abmlog` repository root.

Run the `abmlog` for the CASCADE 2.0 topology:
```
java -cp target/abmlog-1.5.0-jar-with-dependencies.jar eu.druglogics.abmlog.RandomBooleanModelGenerator --file=test/cascade_2_0.sif --num=3000
```

Next, prune the resulting models to only the ones that have 1 stable state (should be $1292$) using the simple bash script [process_models.sh](https://github.com/bblodfon/ags-paper-1/blob/master/scripts/process_models.sh) inside the generated `models` directory from `abmlog`.

```
cd pathTo/druglogics-synergy/ags_cascade_2.0
```

- Move the `abmlog`-generated `models` dir inside the `ags_cascade_2.0` dir
- Use `attractor_tool`: `biolqm_stable_states` in the `config` file
- Use `synergy_method: hsa` or `synergy_method: bliss` in the `config` file (run twice below command, changing the `project` to `cascade_2.0_random_bliss`)

Run Drabme via `druglogics-synergy`:

```
java -cp ../target/synergy-1.2.0-jar-with-dependencies.jar eu.druglogics.drabme.Launcher --project=cascade_2.0_random_hsa --modelsDir=models --drugs=drugpanel --perturbations=perturbations --config=config --modeloutputs=modeloutputs
```

The above procedure is the same for CASCADE 1.0. Changes:

- The `abmlog`-generated `models` directory should be put inside the `ags_cascade_1.0` of `druglogics-synergy`
- The network file is now for the CASCADE 1.0 (`network.sif` inside the `ags_cascade_1.0`)
- The Drabme command should be run with `--project=cascade_1.0_random_hsa` and `--project=cascade_1.0_random_bliss` respectively

# R session info {-}

```{r session info, comment=""}
xfun::session_info()
```

# References {-}
